---
title: "Capstone Final Code"
author: "Miriam Brem"
date: "2026-04-24"
output: html_document
---

# Preliminary Loading 
```{r load packages}
# install.packages("openxlsx", dependencies = TRUE) # Noted here because I did not have it downloaded, so for people reviewing code
library(openxlsx) #Fixes Excel Arrest Date Upload error 
library(readxl) #Loads excel packages onto R
library(tidyverse) #Loads tools like %>%
library(ggplot2) #Loads tool for creating visualizations 
library(readxl) # Loads and combines sentencing data 
library(purrr)  # Loads and combines sentencing data 
```

# Incoming/Outgoing Cases 


## Preparing the Datasets for Analysis 
I was having major issues with my computer memory, so I had to do a lot of filtering and steps in a somewhat illogical order to get the d and cr numbers datasets to merge. First, I am going to load the datasets and select the bare minimum of columns. 
```{r load new data}
# select columns to make the datasets smaller 
cr_data <- read.csv("Data/case_data_cr.csv") %>% 
  select(Party.ID, Case.Number, Received.Date, Unit, Case.Status)
d_data <- read.csv("Data/case_data_d.csv") %>% 
  rename(Party.ID = Praty.ID) %>%  # Fix column name typo
  select(Party.ID, Case.Number, Received.Date, Unit, Case.Status)
cases_closed_data <- read.csv("Data/Cases Closed Data.csv") %>% 
  select(Party.ID, Event.Date)
```
Now I want to merge the first two data sets and then take the data to the case-defendant level
```{r merge cr and d data and go to case defendant level}
date_formats <- c("mdy", "ymd", "dmy", "mdy HM", "ymd HM") # make a list of all of the potential date formats 

# Not enough memory to do the merge right away, so go to case defendant level first
question12_df <- bind_rows(d_data, cr_data)  %>% # Merge the cr and d data 
  # create opening year variable 
  mutate(Received.Date = parse_date_time(Received.Date, orders = date_formats)) %>%
  mutate(Case.Opening.Year = year(Received.Date)) %>%  # Uses Received date for opening year 
  # Go to Case-Defendant level  
  group_by(Case.Number, Case.Opening.Year) %>% 
    slice(1) %>% 
  ungroup() %>% 
  # Filter out a few more units/times
  filter(Unit != "HIDTA") %>% 
  filter(Unit != "SB440") %>%
  filter(Unit != "No Unit Entered") %>%
  filter(Case.Opening.Year >= 2019) # not before 2018
```

Now the data set should be small enough to merge the datasets with my computer memory capacity. So now I am going to merge the cr/d data with the case closed data to get the event date information.
```{r merge cases closed data}
# Now there should be enough memory for the 
question12_df<- left_join(question12_df, cases_closed_data, by = "Party.ID", relationship = "many-to-many") # add the event date where applicable 
```
Now I need to standardize the event date column and create the Case.Closing.Year variable. 

```{r standardize dates}
question12_df  <- question12_df %>% # Parse dates 
  mutate(Event.Date = parse_date_time(Event.Date, orders = date_formats))

question12_df <- question12_df %>% # create closing year 
    mutate(`Case.Closing.Year` = year(`Event.Date`))  %>% # Uses the year that Event/resolution occurred
    filter(Case.Closing.Year >= 2019) %>%
    mutate(`Unit.Type` = if_else(Unit == "Trial Division", "Trial Division", "Specialty Unit")) 
```

Now I am done preparing the data and I can create my visualizations:

```{r two line visualization}
vis_data <- question12_df  %>% # Filter the data a bit 
  filter(!(Event.Date >= as.Date("2024-07-01"))) %>% # not outgoing after 4/25
  filter(!(Received.Date >= as.Date("2024-07-01"))) # not incoming after 4/25

incoming <- vis_data  %>%  
  filter(!is.na(Received.Date)) %>%
  mutate(Period = floor_date(Received.Date, unit = "6 months")) %>%
  group_by(Period) %>%  
  summarise(Count = n(), .groups = "drop") %>%
  mutate(Line = "Incoming")

outgoing <- vis_data %>%
  filter(!is.na(Event.Date)) %>%
  mutate(Period = floor_date(Event.Date, unit = "6 months")) %>%
  group_by(Period) %>%
  summarise(Count = n(), .groups = "drop") %>%
  mutate(Line = "Outgoing")

vis_plot_data <- bind_rows(incoming, outgoing)
class(vis_plot_data$Period)

vis_plot_data$Period <- as.Date(
vis_plot_data$Period, format = "%Y-%m-%d")
# Plot
ggplot(vis_plot_data, aes(x = Period, y = Count, color = Line, group = Line)) +
  geom_line() +
  geom_point() +
  scale_x_date(date_breaks = "6 months", date_labels = "%b %Y") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 35, hjust = 1)) +
  labs(
    title = "Case-Defendant Level Incoming vs Outgoing Cases Over Time by Unit Type",
    x = "Time Period",
    y = "Number of Cases",
    color = "Case Type"
  ) +
  scale_color_manual(values = c(
    "Incoming"= "navy",
    "Outgoing"  = "tan4")) +
  # Add lines for Covid
  geom_vline(xintercept = as.Date("2020-03-13"), color = "red", linetype = "dashed") +
  annotate("text", x = as.Date("2020-03-13"), y = max(vis_plot_data$Count) * 0.93,
           label = "COVID Shutdown", color = "red", angle = 90, vjust = -1, size = 2) +
  geom_vline(xintercept = as.Date("2021-06-21"), color = "red", linetype = "dashed") +
  annotate("text", x = as.Date("2021-06-21"), y = max(vis_plot_data$Count) * 0.95,
           label = "Full Reopening", color = "red", angle = 90, vjust = 2, size = 2) +
  geom_vline(xintercept = as.Date("2021-04-03"), color = "orange", linetype = "dashed") +
  annotate("text", x = as.Date("2021-04-03"), y = max(vis_plot_data$Count) * 0.92,
           label = "Grand Jury Resumes", color = "orange", angle = 90, vjust = -1, size = 2)

rm(vis_plot_data, incoming, outgoing)
```

Now, to get a better idea of the differences between Trial Division and the other Specialty Units (except for HIDTA and SB440), I decided to create the same visualization but separated by the unit type. 

```{r three line incoming and outgoing visualization}
incoming <- vis_data  %>%  
  filter(!is.na(Received.Date)) %>%
  mutate(Period = floor_date(Received.Date, unit = "6 months")) %>%
  group_by(Period) %>%  
  summarise(Count = n(), .groups = "drop") %>%
  mutate(Line = "Incoming (All Units)")

outgoing <- vis_data %>%
  filter(!is.na(Event.Date)) %>%
  mutate(Period = floor_date(Event.Date, unit = "6 months")) %>%
  group_by(Period, Unit.Type) %>%
  summarise(Count = n(), .groups = "drop") %>%
  mutate(Line = paste(Unit.Type, "Outgoing", sep = " - "))

vis_plot_data <- bind_rows(incoming, outgoing)

vis_plot_data$Period <- as.Date(
vis_plot_data$Period, format = "%Y-%m-%d")
# Plot
ggplot(vis_plot_data, aes(x = Period, y = Count, color = Line, group = Line)) +
  geom_line() +
  geom_point() +
  scale_x_date(date_breaks = "6 months", date_labels = "%b %Y") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 35, hjust = 1)) +
  labs(
    title = "Case-Defendant Level Incoming vs Outgoing Cases Over Time by Unit Type",
    x = "Time Period",
    y = "Number of Cases",
    color = "Case Type"
  ) +
  scale_color_manual(values = c(
    "Incoming (All Units)"       = "navy",
    "Trial Division - Outgoing"  = "tan4",
    "Specialty Unit - Outgoing"  = "tan3")) +
  # Add lines for Covid
  geom_vline(xintercept = as.Date("2020-03-13"), color = "red", linetype = "dashed") +
  annotate("text", x = as.Date("2020-03-13"), y = max(vis_plot_data$Count) * 0.93,
           label = "COVID Shutdown", color = "red", angle = 90, vjust = -1, size = 2) +
  geom_vline(xintercept = as.Date("2021-06-21"), color = "red", linetype = "dashed") +
  annotate("text", x = as.Date("2021-06-21"), y = max(vis_plot_data$Count) * 0.95,
           label = "Full Reopening", color = "red", angle = 90, vjust = 2, size = 2) +
  geom_vline(xintercept = as.Date("2021-04-03"), color = "orange", linetype = "dashed") +
  annotate("text", x = as.Date("2021-04-03"), y = max(vis_plot_data$Count) * 0.92,
           label = "Grand Jury Resumes", color = "orange", angle = 90, vjust = -1, size = 2)

rm(vis_plot_data, incoming, outgoing)
```

```{r four line incoming and outgoing visualization}
incoming <- vis_data  %>%  
  filter(!is.na(Received.Date)) %>%
  mutate(Period = floor_date(Received.Date, unit = "6 months")) %>%
  group_by(Period, Unit.Type) %>%  
  summarise(Count = n(), .groups = "drop") %>%
  mutate(Line = paste(Unit.Type, "Incoming", sep = " - "))

outgoing <- vis_data %>%
  filter(!is.na(Event.Date)) %>%
  mutate(Period = floor_date(Event.Date, unit = "6 months")) %>%
  group_by(Period, Unit.Type) %>%
  summarise(Count = n(), .groups = "drop") %>%
  mutate(Line = paste(Unit.Type, "Outgoing", sep = " - "))

vis_plot_data <- bind_rows(incoming, outgoing) %>%
  mutate(Line = factor(Line, levels = c(
    "Trial Division - Outgoing", "Specialty Unit - Outgoing", "Trial Division - Incoming", "Specialty Unit - Incoming")))

vis_plot_data$Period <- as.Date(
vis_plot_data$Period, format = "%Y-%m-%d")
# Plot
ggplot(vis_plot_data, aes(x = Period, y = Count, color = Line, group = Line)) +
  geom_line() +
  geom_point() +
  scale_x_date(date_breaks = "6 months", date_labels = "%b %Y") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 35, hjust = 1)) +
  labs(
    title = "Case-Defendant Level Incoming vs Outgoing Cases Over Time by Unit Type",
    x = "Time Period",
    y = "Number of Cases",
    color = "Case Type"
  ) +
  scale_color_manual(values = c(
    "Trial Division - Incoming" = "navy",
    "Specialty Unit - Incoming" = "royalblue",
    "Trial Division - Outgoing"  = "tan4",
    "Specialty Unit - Outgoing"  = "tan3")) +
  # Add lines for Covid
  geom_vline(xintercept = as.Date("2020-03-13"), color = "red", linetype = "dashed") +
  annotate("text", x = as.Date("2020-03-13"), y = max(vis_plot_data$Count) * 0.93,
           label = "COVID Shutdown", color = "red", angle = 90, vjust = -1, size = 2) +
  geom_vline(xintercept = as.Date("2021-06-21"), color = "red", linetype = "dashed") +
  annotate("text", x = as.Date("2021-06-21"), y = max(vis_plot_data$Count) * 0.95,
           label = "Full Reopening", color = "red", angle = 90, vjust = 2, size = 2) +
  geom_vline(xintercept = as.Date("2021-04-03"), color = "orange", linetype = "dashed") +
  annotate("text", x = as.Date("2021-04-03"), y = max(vis_plot_data$Count) * 0.92,
           label = "Grand Jury Resumes", color = "orange", angle = 90, vjust = -1, size = 2)

#rm(vis_plot_data, incoming, outgoing)
```




------------------------------------------------------------------------------


# Case Outcomes 


## Preparing the Datasets for Analysis 
First, we are going to load the data sets. We are going to combine the `cases closed data` and the `sentencing data`. 

**VIEWER NOTE: This uses an excel sheet for Sentencing Data instead of a CSV, so download accordingly. Put it in a folder and download from there. Also expect this to take a while to load.**
```{r prepare q13 data for analysis}
cases_closed_data <- read.csv("Data/Cases Closed Data.csv") 

# Used for Sentencing Information
files_path <- "/Users/karizhou/Desktop/SPRING 26/Capstone/Data/Sentencing Data" # File path - this changes depending on computer
# Used as a Marker for Trials
trial_data <- read.csv("Data/Trial Data.csv")

# get all of the xlsx files
xlsx_files <- list.files(path = files_path,
                        pattern = "\\.xlsx$",
                        full.names = TRUE)
xlsx_files <- xlsx_files[!grepl("/~\\$", xlsx_files)] # Remove Excel temporary files (~$) - this will not work if you have any of the files open
sentencing_data <- map_dfr(xlsx_files, read_excel) # Combine all of the files into one df 

# Clean the data to remove repeated rows 
sentencing_data <- sentencing_data %>% 
  distinct(`Party ID`, `ChargeID`, .keep_all = TRUE) # Keep one row for each distinct Party ID/Charge ID combo 

sentencing_data_cleaning_subset <- sentencing_data[,c("CaseNbr", "Plea", "Disposition", "Disposition Date")] # Only pick certain rows

question13_df <- cases_closed_data %>%
  left_join(sentencing_data_cleaning_subset, by = c("Case.Number"= "CaseNbr"), relationship = "many-to-many") %>%  # They do not match up because cases-close has more information then sentencing (not all cases receive a sentence) so this will add NULL for sententencing where they do not overlap 
  mutate(Trial = Case.Number %in% trial_data$Case.Number) %>% 
  mutate(year = year(`Disposition Date`))   # Create a year column

rm(sentencing_data_cleaning_subset) # Remove unneeded df
#Utilized AI for this block of code. See AI Use Statement Q11 1 at the end.
```

### Analyzing the Data 
This question asks me for the annual conviction rate of cases. The conviction rate shows how many trials result in a conviction. It is calculated using this formula: 

**Conviction Rate** = Total Number of Convictions/ Total Number of Trials

So, I want to create a function to calculate this for a given dataset.
```{r creating a function to calculate conviction rate}
# Create a list of all dispositions that count as convictions 
  conviction_list <- c("Bench Verdict Guilty (Convicted/Adjudicated)", "First Offender Act - Pending", "First Offender Act Completed", "Guilty", "Jury Verdict Guilty (Convicted/Adjudicated)","First Offender Act Revoked", "First Offender Act Tolled", "First Offender Act Unsatisfactory", "Nolo") 
  
calculate_conviction_df <- function(df) {
  # Find the total number of convictions per year 
  conviction <- df %>% 
    filter(`Disposition` %in% conviction_list) %>% # Filter for only convictions 
    group_by(`Party.ID`, year) %>% 
     slice(1) %>% # Only keep first entry from each defendant group
    ungroup() %>%
    group_by(year) %>% 
    summarize(total_conviction_count = n()) # Count the number of convictions per year 
  
  # Find the total number of trials per year 
  trial <- df %>% 
    filter(Trial == TRUE) %>% # Filter for only trials 
    group_by(`Party.ID`, year) %>% 
      slice(1) %>% # Only keep first entry from each defendant group 
    ungroup() %>%
    group_by(year) %>% 
    summarize(total_trial_count = n()) # Count the number of trial per year 
  
  # Find the conviction rate per year 
    conviction_rate <- full_join(conviction, trial, by = "year") %>% # Combine conviction and trial dfs 
    mutate(conviction_rate = total_conviction_count / total_trial_count) %>% # calculate conviction rate
    select(year, conviction_rate) %>% # clean up final df 
    filter(year >= 2018) %>% 
    arrange(desc(year))
  
  return(conviction_rate)}
```

Note: The exact list of dispositions that shall be counted as convictions were not listed in the same format, so some analysis choices were made regarding what counts as a conviction. 

```{r investigating the number of guilty pleas that do not go to trial}
acquittals_list <- c("Jury Verdict Not Guilty", "Bench Verdict Not Guilty (Convicted/Adjudicated)", "Not Guilty by Reason of Insanity - Restrict", "Not Guilty", "Not Guilty - Restrict", "Jury Verdict Not Guitly by Reason of Insanity")


# Use very similar code as above. 
results_list <- list() # Create list for results 
unit_list <- c("Trial Division", "Specialty Unit")

plot_vis <- question13_df %>% # Create a column with if it belongs to the Trial Division or not 
  filter(Unit != "No Unit Entered") %>% 
  mutate(`Unit.Type` = if_else(Unit == "Trial Division", "Trial Division", "Specialty Unit")) %>% 
  mutate(win = if_else(Disposition %in% conviction_list, "Convicted", "Not Convicted")) %>% 
  mutate(win = case_when(Disposition %in% conviction_list ~ "Convicted", 
                         Disposition %in% acquittals_list ~ "Acquitted" , 
                         TRUE ~ "Dismissed"))  %>%
  group_by(`Party.ID`, year) %>%  # Make into Case-Defendant level 
    arrange(desc(Disposition %in% conviction_list)) %>% # Put convictions at the top of the list - Includes Pleas 
    slice(1) %>% # Only keep first entry from each defendant group
    ungroup()
plot_vis$win <- factor(plot_vis$win, levels = c("Convicted", "Dismissed","Acquitted"))

# Create a loop for each division
for(i in 1:length(unit_list)) { # Run one time for each division. So 2 times 
  current_unit <- unit_list[i] # Keep track of which division we're on 

  # Find and store the count of cases 
  unit_results <- plot_vis %>% 
    filter(`Unit.Type` == current_unit) %>%
    group_by(year, win) %>% # Separate by year and if it was a win 
    summarise(case_count = n()) %>% 
    ungroup() %>% 
    mutate(unit = current_unit)
  
  results_list[[i]] <- unit_results # Store the conviction rate df
}
results <- bind_rows(results_list) # Combine all of the df into one by year
results$win <- factor(results$win, levels = c("Acquitted", "Dismissed", "Convicted"))

# Create visualization
ggplot(results, aes(x = year, y = case_count, fill = win, group = unit)) + 
  geom_col(position = "stack") +
  facet_wrap(~ unit) +
  theme_minimal() +
  labs(title = 'Case Defendant-Level Case Counts for Trial Division vs. Other Units',
      x = "Year", y = "Case Count", fill = "Case Outcome") + 
  scale_x_continuous(breaks = seq(2019, 2025, by = 1)) + # Make sure the x-axis shows every year 
  scale_fill_manual(values = c(
    "Convicted"     = "darkblue",
    "Acquitted" = "black", 
    "Dismissed" = "slategrey")) +
  # Add lines for Covid 
  geom_vline(xintercept = 2020, color = "red", linetype = "dashed") + 
  annotate("text", x = 2020, y = 2100, label = "COVID Shutdown", color = "red", angle = 90, vjust = -1, size = 2) +
  geom_vline(xintercept = 2021, color = "red", linetype = "dashed") + 
  annotate("text", x = 2021, y = 2121, label = "Full Reopening", color = "red", angle = 90, vjust = -1, size = 2)

rm(results, unit_list, unit_results, results_list) # Remove dfs that will not be reused
```

To replace the conviction rate and represent how many cases actually go to trial, here is a similar visualization but only with the cases that went to trial. 
```{r Defendant-Level Trial Conviction Counts of Trial Division vs. Other Units}
# Use very similar code as above. 
results_list <- list() # Create list for results 
unit_list <- c("Trial Division", "Specialty Unit")

plot_vis <- question13_df %>% # Create a column with if it belongs to the Trial Division or not 
  filter(Unit != "No Unit Entered") %>% 
  mutate(`Unit.Type` = if_else(Unit == "Trial Division", "Trial Division", "Specialty Unit")) %>% 
  mutate(win = case_when(Disposition %in% conviction_list ~ "Convicted", 
                         Disposition %in% acquittals_list ~ "Acquitted" , 
                         TRUE ~ "Dismissed"))  %>%
  filter(Trial == TRUE) %>% 
  group_by(`Party.ID`, year) %>%  # Make into Case-Defendant level 
    arrange(desc(Disposition %in% conviction_list)) %>% # Put convictions at the top of the list - Includes Pleas 
    slice(1) %>% # Only keep first entry from each defendant group
    ungroup()

plot_vis$win <- factor(plot_vis$win, levels = c("Convicted", "Dismissed","Acquitted"))


# Create a loop for each division
for(i in 1:length(unit_list)) { # Run one time for each division. So 2 times 
  current_unit <- unit_list[i] # Keep track of which division we're on 

  # Find and store the count of cases 
  unit_results <- plot_vis %>% 
    filter(Unit.Type == current_unit) %>%
    group_by(year, win) %>% # Separate by year and if it was a win 
    summarise(case_count = n()) %>% 
    ungroup() %>% 
    mutate(unit = current_unit)
  
  results_list[[i]] <- unit_results # Store the conviction rate df
}
results <- bind_rows(results_list) # Combine all of the df into one by year
results$win <- factor(results$win, levels = c("Acquitted", "Dismissed", "Convicted"))

# Create visualization
ggplot(results, aes(x = year, y = case_count, fill = win, group = unit)) + 
  geom_col(position = "stack") +
  facet_wrap(~ unit) +
  theme_minimal() +
  labs(title = 'Case Defendant-Level Trial Conviction Counts of Trial Division vs. Other Units',
      x = "Year", y = "Case Count", fill = "Case Outcome") + 
  scale_x_continuous(breaks = seq(2019, 2025, by = 1)) + # Make sure the x-axis shows every year 
  scale_fill_manual(values = c(
    "Convicted"     = "royalblue",
    "Acquitted" = "black", 
    "Dismissed" = "slategrey")) +
  # Add lines for Covid 
  geom_vline(xintercept = 2020, color = "red", linetype = "dashed") + 
  annotate("text", x = 2020, y = 96, label = "COVID Shutdown", color = "red", angle = 90, vjust = -1, size = 2) +
  geom_vline(xintercept = 2021, color = "red", linetype = "dashed") + 
  annotate("text", x = 2021, y = 98, label = "Full Reopening", color = "red", angle = 90, vjust = -1, size = 2)

rm(results, unit_list, unit_results, results_list, plot_vis) # Remove dfs that will not be reused
```

But this doesn't tell the whole story, there are a lot of cases that are achieved without a trial, so this visualization is meant to show that: 
```{r Defendant-Level Non-Trial Guilty Plea Counts for Trial Division vs. Other Units}
# Use very similar code as above. 
results_list <- list() # Create list for results 
unit_list <- c("Trial Division", "Specialty Unit")

plot_vis <- question13_df %>% # Create a column with if it belongs to the Trial Division or not 
  filter(Unit != "No Unit Entered") %>% 
  mutate(`Unit.Type` = if_else(Unit == "Trial Division", "Trial Division", "Specialty Unit")) %>% 
  mutate(win = case_when(Disposition %in% conviction_list ~ "Plead Guilty", 
                         Disposition %in% acquittals_list ~ "Acquitted" , 
                         TRUE ~ "Dismissed"))  %>%
  filter(Trial == FALSE) %>% 
  filter(Trial != TRUE) %>% 
  group_by(`Party.ID`, year) %>%  # Make into Case-Defendant level 
    arrange(desc(Disposition %in% conviction_list)) %>% # Put convictions at the top of the list - Includes Pleas 
    slice(1) %>% # Only keep first entry from each defendant group
    ungroup()

plot_vis$win <- factor(plot_vis$win, levels = c("Plead Guilty", "Dismissed","Acquitted"))

# Create a loop for each division
for(i in 1:length(unit_list)) { # Run one time for each division. So 2 times 
  current_unit <- unit_list[i] # Keep track of which division we're on 

  # Find and store the count of cases 
  unit_results <- plot_vis %>% 
    filter(`Unit.Type` == current_unit) %>%
    group_by(year, win) %>% # Separate by year and if it was a win 
    summarise(case_count = n()) %>% 
    ungroup() %>% 
    mutate(unit = current_unit)
  
  results_list[[i]] <- unit_results # Store the conviction rate df
}
results <- bind_rows(results_list) # Combine all of the df into one by year
results$win <- factor(results$win, levels = c("Acquitted", "Dismissed", "Plead Guilty"))

# Create visualization
ggplot(results, aes(x = year, y = case_count, fill = win, group = unit)) + 
  geom_col(position = "stack") +
  facet_wrap(~ unit) +
  theme_minimal() +
  labs(title = 'Case Defendant Non-Trial Guilty Plea Counts for Trial Division vs. Other Units',
      x = "Year", y = "Case Count", fill = "Case Outcome") + 
  scale_x_continuous(breaks = seq(2019, 2025, by = 1)) + # Make sure the x-axis shows every year 
  scale_fill_manual(values = c(
    "Plead Guilty"= "blue",
    "Acquitted" = "black", 
    "Dismissed" = "slategrey")) +
  # Add lines for Covid 
  geom_vline(xintercept = 2020, color = "red", linetype = "dashed") + 
  annotate("text", x = 2020, y = 2082, label = "COVID Shutdown", color = "red", angle = 90, vjust = -1, size = 2) +
  geom_vline(xintercept = 2021, color = "red", linetype = "dashed") + 
  annotate("text", x = 2021, y = 2115, label = "Full Reopening", color = "red", angle = 90, vjust = -1, size = 2)

rm(results, unit_list, unit_results, results_list, plot_vis) # Remove dfs that will not be reused
```

# SVC 

## Preparing the Datasets for Analysis 
First, we are going to load the data sets. We are loading the 'Plea and Disposition Data', 'Case Attorney Data' and 'Case Sentencing Data'.  

```{r loads Q11 datasets}
# Used to matched cases with their attorney
attorney_data <- read.csv("Data/Case Attorney Data.csv")

# Used for Division Data 
division_data <- read.csv("Data/Case Division Data _ Post Standing Order _ Post-Indictment Assignments(Sheet1).csv")
```

This question asks us to compare attorneys and judges from within the trial division, so we can start by filtering the `attorney data` to only include cases handled by the trial division and only keep the attorneys were never removed and thus closed the trial. 
```{r filter attorney data to only include the trial division}
attorney_data_cleaning <- attorney_data %>%
  filter(`Unit` == "Trial Division") %>% # Filter out any division except for the Trial Division
  filter(`Removed.Date` == "NULL") # Only keep the attorneys were never removed and thus closed the trial
```

This question asks us to compare the different Trial Division attorneys and judges. Thus, to do so, we need to be able to connect each case with it's respective judge. So, we are going to merge the 'Plea and Disposition Data' with the 'Case Attorney Data' and 'Case Division Data _ Post Standing Order _ Post-Indictment Assignments'.
```{r merging plea and attorney datasets}
# Keep only a few columns from the original dataset to be used for merging
sentencing_data_cleaning_subset <- sentencing_data[,c("CaseNbr", "ChargeID", "Offense Code", "Offense","Plea","Disposition", "Disposition Date", "SentenceYears", "Sentence Type")] 
attorney_data_cleaning_subset <- attorney_data_cleaning[,c("Case.Number", "Defendant.Name","Defendant.Party.ID","Attorney.Name", "Role", "Assigned.Date")] 
division_data_cleaning_subset <- division_data[,c("Case.Number", "Division", "Case.ID")]

# Combine datasets
intial_question11_df <- merge(attorney_data_cleaning_subset, sentencing_data_cleaning_subset, by.x = "Case.Number", by.y = "CaseNbr", all.x = TRUE) 

intial_question11_df <- intial_question11_df %>%
  left_join(division_data_cleaning_subset, by = "Case.Number",  relationship = "many-to-many") %>%  # Check that these actually match
  mutate(Trial = Case.Number %in% trial_data$Case.Number)
```

Since I am merging together a CSV and an XLSX dataframe, I want to rename some of the columns for consistency
```{r rename sentencing data columns}
intial_question11_df <- rename(intial_question11_df,
                                Offense.Code = `Offense Code`, 
                                Dispostion.Date = `Disposition Date`, 
                                Sentence.Years = SentenceYears, 
                                Sentence.Type = `Sentence Type`)
```


Now that I've created the primary dataset that I plan to use for this question I am going to remove some of the intermediary steps. For future questions, I'm probably going to have to adjust and use these datasets again, so I am going to remove them from my environment to prevent future confusion. 
```{r removing unneeded dataframes}
# Removes the datasets from the environment
rm(attorney_data_cleaning_subset, attorney_data_cleaning, division_data_cleaning_subset, sentencing_data_cleaning_subset)
```

Now, since Q11 asks about "defendants charged initially with serious violent crime", I am going to filter the 'Plea and Disposition Data' so that it only includes these defendants. 

To do so, I am going to group by party ID and the filter so that only defendants with one of the following predefined violent crimes remain: 

| Offense | Offense Code | 
| :--- | :---: |
| Armed Robbery |16-8-41|
| Kidnapping | 16-5-40 |
| Aggravated Assault | 16-5-21 |
| Aggravated Battery | 16-5-24 |
| Trafficking | 16-13-31 |
| Hijacking a Motor Vehicle | 16-5-44 |

```{r filter for an initial charging of a serious violent crime}
# Create a list containing all of the offense codes of interest 
offense_codes <- c("16-8-41","16-5-40","16-5-21","16-5-24","16-13-31","16-5-44")
offense_code_w_endings <- paste0("^(", paste(offense_codes, collapse="|"), ")($|[^0-9])") # Keeps offense codes in order, allows for different offense code endings, and makes sure that there are no additional code numbers on the end

# Group by defendant and filter out irrelevant defendants  
question11_df <- intial_question11_df %>%
  mutate(Crime = case_when( # Create a column with crime type
      grepl("^16-8-41",  Offense.Code) ~ "Armed Robbery",
      grepl("^16-5-40",  Offense.Code) ~ "Kidnapping",
      grepl("^16-5-21",  Offense.Code) ~ "Aggravated Assault",
      grepl("^16-5-24",  Offense.Code) ~ "Aggravated Battery",
      grepl("^16-13-31", Offense.Code) ~ "Drug Trafficking",
      grepl("^16-5-44",  Offense.Code) ~ "Hijacking a Motor Vehicle")) %>%
  group_by(`Defendant.Party.ID`) %>% # Group by party ID
  filter(any(grepl(offense_code_w_endings, `Offense.Code`))) %>% #Gets rid of groups without any matching Offense Codes
  ungroup() #Stops grouping my offender 

rm(intial_question11_df)
#Utilized AI for this block of code. See AI Use Statement Q11 2 at the end.
```

## Analyzing the Data

Now, before I can look at anything division specific, I want to see the number of SVCs per division. 
```{r Counts visualization of sentence type by division }
division_list <- c("Division 1","Division 2","Division 3","Division 4","Division 5",
                   "Division 6","Division 7","Division 8","Division 9", "Division 10")


sentence_summary <- question11_df %>% # Prepare data by making it defendant-level by greatest charge 
  filter(Division %in% division_list) %>% # remove rows with no division
  filter(`Sentence.Type` != "NULL") %>% # Remove rows with no Sentence Type
  filter(!is.na(Crime)) %>% # Remove extraneous crimes 
  distinct(Defendant.Party.ID, Case.ID, Crime, .keep_all = TRUE) %>% # Only keep the SVCs
  ungroup() 
  
count <- sentence_summary %>% 
  group_by(Division, Sentence.Type) %>% 
  summarize(total_case_count = n(), .groups = "drop") %>% 
  ungroup()

count$Sentence.Type <- factor(count$`Sentence.Type`, levels = c("House Arrest", "Suspended Sentence","Probation", "Confinement", "Confinement/Probation"))
count$Division <- factor(count$Division, levels = c("Division 1","Division 2","Division 3","Division 4","Division 5","Division 6","Division 7","Division 8","Division 9", "Division 10"))

# Create visualization
ggplot(count, aes(x = Division, y = total_case_count, fill = Sentence.Type)) +
  geom_bar(position = "stack", stat = "identity") +                           
  theme_minimal() +
  theme(legend.position = "right",axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    title = "Serious Violent Crimes Charge Sentencing Types by Trial Division",
    x = "Division",
    y = "Number of Sentences",
    fill = "Sentence Type") + 
  scale_fill_manual(values = c(
    "Confinement" = "cornflowerblue",
    "Confinement/Probation" = "royalblue3",
    "House Arrest" = "tan4",  
    "Probation" = "orange3"))

rm(sentence_summary, count)
```

---------------------------------------------------------------------------------------

# Appendix 


## Statistical Testing 
Now I want to use statistical testing to see what 
```{r outgoing cases stats testing}
annual_count <- list() # Create list for number of cases per crime 
annual_list <- c("2019", "2020", "2021", "2022","2023","2024","2025")

outgoing_sig_testing <- question12_df %>%
  filter(!is.na(Event.Date)) %>%
  mutate(Period = floor_date(Event.Date, unit = "3 months")) %>%
  group_by(Period, Unit.Type) %>%
  summarise(outgoing_count = n(), .groups = "drop")

# Create time periods
pre_covid <- outgoing_sig_testing %>%
  filter(Period < as.Date("2020-03-13"))
covid <- outgoing_sig_testing %>%
  filter(Period >= as.Date("2020-03-13") & Period < as.Date("2021-06-21"))
post_covid <- outgoing_sig_testing %>%
  filter(Period >= as.Date("2021-06-21"))

# Pre COVID t-test 
special_outgoing_counts <- pre_covid %>%  # Special Units 
  filter(Unit.Type == "Specialty Unit") %>%  
  pull(outgoing_count)

trial_outgoing_counts <- pre_covid %>%  # Trial Division
  filter(Unit.Type == "Trial Division") %>%  
  pull(outgoing_count)

t.test(special_outgoing_counts,trial_outgoing_counts)

# COVID t-test 
special_outgoing_counts <- covid %>%  # Special Units 
  filter(Unit.Type == "Specialty Unit")  %>%  
  pull(outgoing_count)

trial_outgoing_counts <- covid %>%  # Trial Division
  filter(Unit.Type == "Trial Division") %>%  
  pull(outgoing_count)


t.test(special_outgoing_counts,trial_outgoing_counts)

# Post-COVID t-test
special_outgoing_counts <- post_covid %>%  # Special Units 
  filter(Unit.Type == "Specialty Unit")  %>%  
  pull(outgoing_count)

trial_outgoing_counts <- post_covid %>%  # Trial Division
  filter(Unit.Type == "Trial Division") %>%  
  pull(outgoing_count)


t.test(special_outgoing_counts,trial_outgoing_counts)
```
Thus, the difference between the count of outgoing cases for Special Units vs. Trial Division is significant for the pre-covid and post-covid periods. It is not significant for the during covid period, but this is likely due to the short nature of the shutdown period. 

I would like to test to see if there is a statistically significant difference between the annual number of ALL Trial Division cases and the annual number of ALL special unit cases:
```{r significance testing between groups all}
annual_count <- list() # Create list for number of cases per crime 
annual_list <- c("2019", "2020", "2021", "2022","2023","2024","2025")

 # Prepare for significance testing 
sig_testing <- question13_df %>% # Create a column with if it belongs to the Trial Division or not 
  filter(Unit != "No Unit Entered") %>% 
  mutate(`Unit.Type` = if_else(Unit == "Trial Division", "Trial Division", "Specialty Unit")) %>% 
  group_by(`Party.ID`, year) %>%  # Make into Case-Defendant level 
    slice(1) %>% # Order doesn't matter, I am only looking at the count 
    ungroup()

# Trial Division 
trial_division_annual_counts <- sig_testing  %>%  
  filter(Unit.Type == "Trial Division")
for(i in 1:length(annual_list)) { # Run one time for each year 
  current_year <- annual_list[i] 
  year_count <-  trial_division_annual_counts %>%  # Create a placeholder to further filter data 
    filter(year == current_year)  %>% # Filter for the year of interest
    summarize(total_case_count = n()) # Count the annual years
  year_count$year <- current_year # Add column to temp df with division and store df 
  annual_count[[i]] <- year_count
  } 
results_annual <- bind_rows(annual_count) # Combine all of the dfs into one using "year" 
testing_trial_division_data <- results_annual$total_case_count

# Trial Division
special_unit_annual_counts <- sig_testing  %>%  
  filter(Unit.Type == "Specialty Unit")

for(i in 1:length(annual_list)) { # Run one time for each year 
  current_year <- annual_list[i] 
  year_count <-  special_unit_annual_counts %>%  # Create a placeholder to further filter data 
    filter(year == current_year)  %>% # Filter for the year of interest
    summarize(total_case_count = n()) # Count the annual years
  year_count$year <- current_year # Add column to temp df with division and store df 
  annual_count[[i]] <- year_count
  } 
results_annual <- bind_rows(annual_count) # Combine all of the dfs into one using "year" 
testing_special_unit_data <- results_annual$total_case_count

# Run the t-test 
t.test(testing_trial_division_data , testing_special_unit_data )
```
As I can see, I am 95% confident that the Trial Division handles between about 220 and 1558 more cases per year than Specialty Units. But, since the number of years is limited, the variation between years could slightly impact the reliability. 

I would like to test to see if there is a statistically significant difference between the annual number of TRIED Trial Division cases and the annual number TRIED of special unit cases:
```{r significance testing between groups all}
annual_count <- list() # Create list for number of cases per crime 
annual_list <- c("2019", "2020", "2021", "2022","2023","2024","2025")

 # Prepare for significance testing 
sig_testing <- question13_df %>% # Create a column with if it belongs to the Trial Division or not 
  filter(Unit != "No Unit Entered") %>% 
  filter(Trial == TRUE) %>% 
  mutate(`Unit.Type` = if_else(Unit == "Trial Division", "Trial Division", "Specialty Unit")) %>% 
  group_by(`Party.ID`, year) %>%  # Make into Case-Defendant level 
    slice(1) %>% # Order doesn't matter, I am only looking at the count 
    ungroup() %>%
mutate(convicted_binary = if_else(Disposition %in% conviction_list, 1L, 0L)) # Mark for convictions 


# Special Units 
special_convict_annual_counts <- sig_testing  %>%  
  filter(Unit.Type == "Specialty Unit")  %>%  
  pull(convicted_binary)

# Trial Division
trial_convict_annual_counts <- sig_testing  %>%  
  filter(Unit.Type == "Trial Division") %>%  
  pull(convicted_binary)


# Run the t-test 
t.test(special_convict_annual_counts, trial_convict_annual_counts)

rm(trial_convict_annual_counts, special_convict_annual_counts, sig_testing)
```
As I can see, this does not have a statistically significant difference.


# Additional Figures 
Now that the data has been prepared, we want to look at the **sentencing outcomes**. The first way that we are going to measure this is through a visualization of the number of convictions made by each judge between the years 2018 and 2025

```{r visualize convictions per judge}
# Lists 
results_list <- list() # Create list for results 
conviction_list <- c("Bench Verdict Guilty (Convicted/Adjudicated)", "First Offender Act - Pending", "First Offender Act Completed", "Guilty", "Jury Verdict Guilty (Convicted/Adjudicated)","First Offender Act Revoked", "First Offender Act Tolled", "First Offender Act Unsatisfactory")

division_list <- c("Division 1","Division 2","Division 3","Division 4","Division 5",
                   "Division 6","Division 7","Division 8","Division 9", "Division 10")

# Create a loop for each division
for(i in 1:length(division_list)) { # Run one time for each divison. So 10 times 
  current_division <- division_list[i] # Keep track of which division we're on 
  
  conviction_count <- question11_df %>%  # Create a placeholder to further filter data 
    filter(Division == current_division, Disposition %in% conviction_list) %>% # Filter for the division of interest
    group_by(Defendant.Party.ID, Dispostion.Date) %>%  # Group by Party ID and Time
      slice(1) %>% # Only keep first entry from each defendant/time group
    summarize(total_conviction_count = n()) # Count the number of convictions per year
  
  # Find and store the conviction rates 
  conviction_count$division <- current_division # Add column to conviction rate df with division
  results_list[[i]] <- conviction_count} # Store the conviction rate df

results <- bind_rows(results_list) # Combine all of the df into one by year 
results$division <- factor(results$division, levels = division_list)

# Create visualization
ggplot(results, aes(x = division, y = total_conviction_count, fill = division)) + 
  geom_col() +
  theme_minimal() +
   theme(legend.position = "right",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Total Defendant-Level Conviction Counts of Serious Violent Crimes by Trial Division",
      x = "Division",
      y = "Total Number of Convictions",
      fill = "Division") + 
  scale_fill_manual(values = c(
    "Division 1" = "darkblue",
    "Division 2" = "royalblue3",
    "Division 3" = "cornflowerblue",
    "Division 4" = "orchid4", 
    "Division 5" = "sienna", 
    "Division 6" = "tan4",  
    "Division 7" = "orange3",
    "Division 8" = "orange",
    "Division 9" = "wheat",
    "Division 10" = "slategrey")) 

rm(results,results_list, conviction_count) # Remove dfs that will not be reused 
```
**Note:** Since it didn't perfectly match the information provided in the brief, choices were made regarding what should qualify as a conviction. 

To build on the Serious Violent Crimes Charge Sentencing Types by Trial Division visualization, I am going to look at the sentencing length by division and by crime type in the three following ways: 

First I need to create a filter it down to people who went to prison and then create a new variable for the length of time that someone was imprisoned.
```{r filter for imprisoned and create a sentence length categories}
division_list <- c("Division 1","Division 2","Division 3","Division 4","Division 5",
                   "Division 6","Division 7","Division 8","Division 9", "Division 10")

# Create a new column with the sentencing time 
sentencing_times <- question11_df %>% 
  filter(Division %in% division_list) %>% # remove rows with no division
  filter(`Sentence.Type` == "Confinement"| `Sentence.Type` == "Confinement/Probation") %>% # Remove rows with no Sentence Type
  filter(!is.na(Crime)) %>% # Remove extraneous crimes 
  distinct(Defendant.Party.ID, Case.ID, Crime, .keep_all = TRUE) %>% # Only keep the SVCs
  arrange(desc(Sentence.Years)) %>%
  ungroup() %>% 
  mutate(Sentence.Years = suppressWarnings(as.numeric(Sentence.Years)))  %>% # Make numeric 
  mutate(Sentence.Time = case_when( 
          is.na(`Sentence.Years`) ~ "Less than One Year",
          Sentence.Years >= 5 ~ "Over Five Years",
          Sentence.Years >= 1 ~ "One to Five Years", 
          Sentence.Years == "NULL" ~ "Less than One Year",
          Sentence.Years == "NA" ~ "Less than One Year")) %>% 
  filter(!is.na(Sentence.Time))  # Get rid of remaining NAs 

# Order levels 
sentencing_times$`Sentence.Time` <- factor(sentencing_times$`Sentence.Time`, levels = c("Over Five Years", "One to Five Years", "Less than One Year"))
sentencing_times$Division <- factor(sentencing_times$Division, levels = c("Division 1","Division 2","Division 3","Division 4","Division 5","Division 6","Division 7","Division 8","Division 9", "Division 10"))
```

Now I can create my first visualization: 
```{r Counts SVC Crime Sentencing Time by Trial Division}
count <- sentencing_times %>% 
  group_by(Division, Sentence.Time) %>% 
  summarize(total_case_count = n(), .groups = "drop") %>% 
  ungroup()

# Create visualization 
ggplot(count, aes(x = Division, y = total_case_count, fill = Sentence.Time)) +
  geom_bar(position = "stack", stat = "identity") +                           
  theme_minimal() +
  theme(legend.position = "right", axis.text.x = element_text(angle = 35, hjust = 1)) +
  labs(
    title = "Charge-Level Serious Violent Crime Sentencing Time by Trial Division",
    x = "Division",
    y = "Count of Sentences",
    fill = "Sentence Time") + 
  scale_fill_manual(values = c(
    "Less than One Year" = "royalblue3",
    "One to Five Years" = "darkblue",  
    "Over Five Years" = "tan4"))
rm(count)
```

Now, I want to specifically look at the sentencing time by crime: 
```{r Counts SVC Crime Sentencing Time by Trial Division}
count <- sentencing_times %>% 
  group_by(Crime, Sentence.Time) %>% 
  summarize(total_case_count = n(), .groups = "drop") %>% 
  ungroup()

# Create visualization 
ggplot(count, aes(x = Crime, y = total_case_count, fill = Sentence.Time)) +
  geom_bar(position = "stack", stat = "identity") +                           
  theme_minimal() +
  theme(legend.position = "right", axis.text.x = element_text(angle = 35, hjust = 1)) +
  labs(
    title = "Charge-Level Serious Violent Crime Sentencing Time by Crime",
    x = "Crime Type",
    y = "Count of Sentences",
    fill = "Sentence Time") + 
  scale_fill_manual(values = c(
    "Less than One Year" = "royalblue3",
    "One to Five Years" = "darkblue",  
    "Over Five Years" = "tan4"))
rm(count)
```

Finally, I can verify that there is enough data and then faccett this visualization by crime: 
```{r Charge counts of SVC Sentencing Time by Division}
sufficent_crime_list <- c("Aggravated Assault", "Armed Robbery")
crimed_data <- sentencing_times %>% 
  filter(Crime %in% sufficent_crime_list, Division != "Division 2") %>%
  group_by(Crime, Sentence.Time, Division) %>% 
  summarize(total_case_count = n(), .groups = "drop") %>% 
  ungroup()

# Create visualization 
ggplot(crimed_data, aes(x = Division, y = total_case_count, fill = Sentence.Time)) +
  geom_bar(position = "stack", stat = "identity") +                           
  theme_minimal() +
  facet_wrap(~ Crime, strip.position = "bottom") +
  theme(legend.position = "right", axis.text.x = element_text(angle = 35, hjust = 1)) +
  labs(
    title = "Charge-Level Serious Violent Crime Sentencing Time by Division and Crime",
    x = "Division",
    y = "Number of Sentences", 
    fill = "Sentence Time") + 
  scale_fill_manual(values = c(
    "Less than One Year" = "royalblue3",
    "One to Five Years" = "darkblue",  
    "Over Five Years" = "tan4"))

rm(sufficent_crime_list, crimed_data, sentencing_times)
```
Additionally, for the last visualization in this chunk, I created a table with the number of charges per division/crime and I only included crimes with over 5 charges in each division in the final version of the visualization.  

Next I want to look at the how sentencing varies depending on whether or not people plea. To do this, I am only going to look at the cases that went to trial and within that, I want to create categories for each defendant saying that they “Plead All Charges”, “Plead Some Charges”, or “Did not Plea”. and I will compare the sentencing across these categories. 

I am going to filter the data to create a column for whether a defendant plead for all, some, or none of their charges. Also, I want to filter it down to the Case-Defendant level. 

```{r}
temp_df <- question11_df %>%
    filter(Trial == TRUE) %>% # Only look at cases that went to trial 
    group_by(Defendant.Party.ID) %>%
    mutate(total_charges = n(), plea_count = sum(Plea == "Guilty" | # Count number of charges/pleas per group
           Plea == "Guilty of Lesser Included Offense"| 
           Plea == "Guilty but Mentally Ill"| 
           Plea == "Guilty-Alford"),
    Plea.Status = case_when(plea_count == 0 ~ "Did not Plea Guilty", 
                  plea_count == total_charges ~ "Plead Guilty for All Charges", 
                  plea_count < total_charges ~ "Plead Guilty for Some Charges")) %>% # Categorize Pleas
  ungroup()
temp_df$Plea.Status <- factor(temp_df$Plea.Status, levels = c("Plead Guilty for All Charges", "Plead Guilty for Some Charges", "Did not Plea Guilty")) 

plot_vis <- temp_df %>% # create a new df that gets rid of repeated charges
  group_by(Defendant.Party.ID, Crime) %>%
  filter(!is.na(Crime)) %>% # Remove extraneous crimes 
  distinct(Defendant.Party.ID, Crime, Plea.Status, .keep_all = TRUE)  # One row per defendant per crime type

count <- plot_vis  %>% 
  group_by(Crime, Plea.Status, Division) %>% 
  summarize(total_case_count = n(), .groups = "drop") %>% 
  ungroup()

# Plot 
ggplot(count, aes(x = Crime, y = total_case_count, fill = Plea.Status)) +
  geom_bar(position ="stack", stat = "identity") +                           
  theme_minimal() +
  theme(legend.position = "right", axis.text.x = element_text(angle = 35, hjust = 1)) +
  labs(title = "Defendant-Level Pleas by Serious Violent Crime Type",
        x = "Crime Type",
        y = "Proportion of Pleas",
        fill = "Plea Type") + 
  scale_fill_manual(values = c(
                  "Did not Plea Guilty" = "orange4", 
                  "Plead Guilty for All Charges"= "orchid4", 
                  "Plead Guilty for Some Charges" = "tan3"))
rm(plot_vis, temp_df, count)
```
I tried to investigate the differences in sentencing through a full trial process, but there were only 16 case defendants that fit the criteria (SVCs that were taken to trial and sentenced) thus I did not investigate further. 

# AI USAGE 

I used AI Tools in the following ways: 
**Tool used:** ChatGPT
**Purpose:** Helped to create a visualization.
**Prompt Used:**
  "Help me program in R. I am working with a dataset, that has a line for each purchase.  First I want to group all of the data together by customer ID and then I want to filter out the customers who did not buy something with any of the product IDs in a list" and later "Change the code so that it includes product ids with something at the end that is not a number"
**Tasks Preformed independently:** I adjusted the variable titles, added descriptive comments, and filtered out repeated rows. 

I used AI Tools in the following ways: 
**Tool used:** Claude AI 
**Purpose:** Assisting with basic data filtering and trouble shooting different `Offense Code` endings.
**Prompt Used:**
  "I am trying to import several excel files into the same data frame in r. How would I do this?"
**Tasks Preformed independently:** I adjusted the file path, added descriptive comments, and filtered the data to remove duplicates.  

I used AI Tools in the following ways: 
**Tool used:** Claude AI 
**Purpose:** Troubleshooting data parsing issues  
**Prompt Used:**
  "In R, I am trying to parse dates from an excel file that I turned into a CSV file. My current strategy works for some rows but not others. Help me fic this."
**Tasks Preformed independently:**  I adjusted the labels to fit the data and then expanded it for all of the relevant columns. 
  
I used AI Tools in the following ways: 
**Tool used:** Claude AI 
**Purpose:** Assisting basic visualization creation. 
**Prompt Used:**
  "In R, I have a function that creates a table with 8 years and success rates for each year. It has two columns, year and success_rate. I want to take a that function and run it 10 times for each of 10 types of success and then use ggplot2 to create a visualization foreach one of the people's annual success rates. Help me code this please", "I want them all on the same graph", "No only run it once per person, but show the data for each year", and "It is only showing me the final person. What am I messing up?"
**Tasks Preformed independently:**  I adjusted variable titles, corrected the code, added descriptive comments, adjusted visualization formatting, and rewrote portions of the code for accuracy, and removed leftover dataframes.

I used AI Tools in the following ways: 
**Tool used:** ChatGPT
**Purpose:** Helped to create a visualization.
**Prompt Used:**
  "Help me program in R. I am working with a dataset, that has a line for each purchase.  First I want to group all of the data together by customer ID and then I want to filter out the customers who did not buy something with any of the product IDs in a list" and later "Change the code so that it includes product ids with something at the end that is not a number"
**Tasks Preformed independently:** I adjusted the variable titles, added descriptive comments, and filtered out repeated rows. 













