---
title: "Question 1"
author: "Kari Zhou"
date: "2026-03-16"
output: html_document
---

######## changes Apr7-14

```{r}
library(readxl)
library(tidyr)
library(dplyr)
library(lubridate)
library(ggplot2)
library(stringr)
library(zoo)
post_standing = read.csv("Data/Case Division Data _ Post Standing Order _ Post-Indictment Assignments(Sheet1).csv")[,-c(10:11)] 
pre_standing = read.csv("Data/Case Division Data _ Pre Standing Order _ Post-Indictment Assignments.csv")
# check column names for rbind later
colnames(post_standing)
colnames(pre_standing)
# data processing, convert date and add stage flag
pre_standing = pre_standing %>% 
  filter(grepl("Division", Division)) %>% 
  mutate(stage = "pre_standing",
         case_nbr = Case.Number,
         party_id = Party.ID,
         case_id = Case.ID,
         defendant = Defendant.Name,
         Division = str_remove(Division, "ivision "),
         date_assigned = as.Date(ifelse(Division.Assignment.TimeStamp.Created == "" | is.na(Division.Assignment.TimeStamp.Created),
                                       NA, Division.Assignment.TimeStamp.Created), 
                                format = "%m/%d/%y"),
         ) %>% 
  
  select(case_nbr,party_id, case_id, defendant, Division, date_assigned, stage ) #20718
  
post_standing = post_standing %>%  
  filter(grepl("Division", Division)) %>% 
  mutate(stage = "post_standing",
         case_nbr = Case.Number,
         party_id = Party.ID,
         case_id = Case.ID,
         defendant = Defendant.Name,
         Division = str_remove(Division, "ivision "),
         date_assigned = as.Date(ifelse(Division.Assignment.TimeStamp.Created == "" | is.na(Division.Assignment.TimeStamp.Created), NA, Division.Assignment.TimeStamp.Created),  
                                 format = "%m/%d/%Y"),
         ) %>% 
  select(case_nbr,party_id, case_id, defendant, Division, date_assigned, stage ) #14241

attorney_levels <- paste(
  rep(paste0("D", 1:10), each = 2),
  rep(c("odd", "even"), times = 10)
)

# define covid shutdown period
covid_period <- data.frame(
  x1 = as.yearqtr("2020-03-01",  format = "%Y-%m-%d"),
  x2 = as.yearqtr("2021-06-01", format = "%Y-%m-%d"),
  y1 = - Inf,
  y2 = Inf
)


```
# link each assignment row with a case received date (earliest)
```{r}
CR_case_data <- read.csv("data/case_data_cr.csv")
CR_case_data <- CR_case_data %>% 
  mutate(
         case_nbr = Case.Number,
         party_id = Party.ID,
         DA_id = DA.ID,
         defendant = Defendant,
         offense_date = as.Date(ifelse(Offense.Date == "" | is.na(Offense.Date),
                                       NA, Offense.Date), 
                                format = "%m/%d/%Y"),
         arrest_date = as.Date(ifelse(Arrest.Date == "" | is.na(Arrest.Date),
                                       NA, Arrest.Date), 
                                format = "%m/%d/%Y"),
         received_date = as.Date(ifelse(Received.Date == "" | is.na(Received.Date),
                                       NA, Received.Date), 
                                format = "%m/%d/%Y"),
         case_status_date = as.Date(ifelse(Case.Status.Date == "" | is.na(Case.Status.Date),
                                       NA, Case.Status.Date), 
                                format = "%m/%d/%Y"),
         offense_code = Offense.Code
         ) %>% 
  select(case_nbr, party_id, DA_id, defendant, offense_date, arrest_date, received_date
         , case_status_date, Offense, offense_code, Unit)

colnames(CR_case_data)  
```

```{r}
# for post standing cases, look for first attorney assignment
# keep the earliest assignment of attorney
# on case-defendant level
attorney_parity_new <- function(df) {
  df_early <- df %>%
    mutate(
      parity = ifelse(
        as.numeric(substr(case_nbr, nchar(case_nbr), nchar(case_nbr))) %% 2 == 0,
        "even",
        "odd"
      ),
      Attorney = paste(Division, parity)
    )
  return(df_early)
}

post_standing_cd = attorney_parity_new(post_standing) %>% filter(!case_nbr == "TESTSTCRIM")
# 10357 cases (post indictment, after standing order), excluding test criminal
pre_standing_cd = attorney_parity_new(pre_standing) %>% filter(!is.na(parity))
# 16665 cases (post indictment, before standing order), excluding case numbers not ending with a number
```


```{r}
post_indictemnt_cd = rbind(post_standing_cd, pre_standing_cd) %>% 
  filter(!is.na(case_nbr)) %>% 
  group_by(case_nbr, party_id
           ) %>% 
  slice(which.min(date_assigned)) # 26381 cases

post_indictemnt_cd$Attorney <- factor(
    post_indictemnt_cd$Attorney,
    levels = attorney_levels
)

post_indictment_all_cd = left_join(post_indictemnt_cd[,c(1:9)], 
                               CR_case_data[,c("received_date", "case_nbr", "party_id", "Unit", "offense_code", "Offense")],
                               by = c("case_nbr", "party_id"))
post_indictment_all_cd = post_indictment_all_cd[!duplicated(post_indictment_all_cd),] 

# for each case-defendant, keep the earliest assigned date
post_indictment_all_cd <- post_indictment_all_cd %>% 
  group_by(case_nbr, party_id) %>% 
  slice(which.min(date_assigned)) %>% 
  filter(!is.na(offense_code))
  #28923, case-defendant level
```

```{r}
# filter for special units 
special_labels = c("DVSA", "CCU", "CAEPAD", "HTIC")

post_indictment_special_units = post_indictment_all_cd %>% 
  filter(Unit %in% special_labels) #3079

# check there's only one row kept for each case-offense-defendant group. 
post_indictment_special_units %>% 
  group_by(case_nbr, offense_code, party_id) %>% 
  filter(n()>1) #0 
```

# major crimes

```{r}
major_crime_levels = c("Homicide/Gangs", "FVPU")

post_indictment_major = post_indictment_all_cd %>% filter(Unit %in% major_crime_levels)

post_indictment_major

```


```{r}
post_indictment_all_cd <- post_indictment_all_cd %>% 
  mutate(Group = case_when(Unit %in% major_crime_levels ~ "Major Crimes",
                           Unit %in% special_labels ~"SVU Crimes",
                           Unit %in% "Trial Division" ~ "Trial Division",
                           TRUE  ~ "No Units Entered"
                           )) %>%  #28923 %>% 
filter(date_assigned > "2017-12-31", date_assigned < "2025-04-01") 


post_indictment_all_cd # 29822, from 2018 to 2025Q1
```

# distribution, assigned date
```{r}
assigned_date <- post_indictment_all_cd %>% 
  filter(!Group == "No Units Entered") %>% 
  mutate(quarter = as.yearqtr(date_assigned, format = "%Y-%m-%d")) %>% 
  group_by(quarter, Group) %>% 
  summarise(count = n()) %>% 
  ggplot(aes(x = quarter, y = count, color = Group))+
  scale_x_yearqtr(format = "%YQ%q", n = 14) +
  geom_line(linewidth = 1)+
    scale_color_manual(values = c(
    "Trial Division" = "navy",
    #"No Units Entered" = "lightblue",
    "Major Crimes" = "tan4",
    "SVU Crimes" = "blue"
  ))+
  geom_rect(data= covid_period, aes(xmin = x1, xmax = x2, ymin = y1, ymax = y2), 
            fill = "orange",
            alpha = 0.3, inherit.aes = FALSE)+
  labs(title = "CR-Cases Assigned to Different Units Over Time (Case-Defendant Level)",
       x = "Time assigned",
       y = "Number of Cases")+
  theme_bw()+
    theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.text = element_text(size = 7),   
    legend.title = element_text(size = 8),
    legend.position = "right"              
  )

assigned_date
```

# distribution, received date
```{r}

received_date <- post_indictment_all_cd %>% 
  filter(!Group == "No Units Entered") %>% 
  mutate(quarter = as.yearqtr(received_date, format = "%Y-%m-%d")) %>% 
  filter(received_date >= "2018-01-01") %>% 
  group_by(quarter, Group) %>% 
  summarise(count = n()) %>% 
  ggplot(aes(x = quarter, y = count, color = Group))+
  scale_x_yearqtr(format = "%YQ%q", n = 14) +
  geom_line(linewidth = 1)+
    scale_color_manual(values = c(
    "Trial Division" = "navy",
    #"No Units Entered" = "lightblue",
    "Major Crimes" = "tan4",
    "SVU Crimes" = "blue"
  ))+
  geom_rect(data= covid_period, aes(xmin = x1, xmax = x2, ymin = y1, ymax = y2), 
            fill = "orange",
            alpha = 0.3, inherit.aes = FALSE)+
  labs(title = "CR-Cases Received by Different Units Over Time (Case-Defendant Level)",
       x = "Time received",
       y = "Number of Cases")+
  theme_bw()+
    theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.text = element_text(size = 7),   
    legend.title = element_text(size = 8),
    legend.position = "right"              
  )

received_date
```

# average per attorney, assigned date
```{r}
post_indictment_all_cd %>% 
   mutate(quarter = as.yearqtr(date_assigned, format = "%Y-%m-%d")) %>% 
  group_by(quarter, Attorney) %>% 
  summarise(count = n()
            ) %>% 
  group_by(quarter) %>% 
  summarise(average = mean(count)) %>% 
  ggplot(aes(x = quarter, y = average))+
  scale_x_yearqtr(format = "%YQ%q", n = 15) +
  geom_line(color = "navy", linewidth = 1)+
  geom_rect(data= covid_period, aes(xmin = x1, xmax = x2, ymin = y1, ymax = y2), 
            fill = "orange",
            alpha = 0.3, inherit.aes = FALSE)+
  labs(title = "CR Cases Assigned Per Attorney Over Time(Case-Defendant Level)",
       x = "Time assigned",
       y = "Quarterly Average")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```
```{r}
post_indictment_all_cd %>% 
   mutate(quarter = as.yearqtr(received_date, format = "%Y-%m-%d")) %>% 
  filter(received_date >= "2018-01-01") %>% 
  group_by(quarter, Attorney) %>% 
  summarise(count = n()
            ) %>% 
  group_by(quarter) %>% 
  summarise(average = mean(count)) %>% 
  ggplot(aes(x = quarter, y = average))+
  scale_x_yearqtr(format = "%YQ%q", n = 15) +
  geom_line(color = "navy", linewidth = 1)+
  geom_rect(data= covid_period, aes(xmin = x1, xmax = x2, ymin = y1, ymax = y2), 
            fill = "orange",
            alpha = 0.3, inherit.aes = FALSE)+
  labs(title = "CR Cases Received by Each Attorney Over Time(Case-Defendant Level)",
       x = "Time received",
       y = "Quarterly Average")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

# side by side graph 
```{r}
gridExtra::grid.arrange(received_date, assigned_date)
```


```{r}

# sum of caseload per attorney, quarterly
attorney_quarter_count  = post_indictment_cr %>% 
  group_by(quarter, Attorney, Group) %>%
  summarise(count = n())

attorney_quarter_count$Attorney <- factor(
    attorney_quarter_count$Attorney ,
    levels = attorney_levels
) 
# average number of cases assigned to each attorney within the quarter. 
attorney_quarter_sum <- attorney_quarter_count %>% group_by(quarter, Group) %>% summarise(average = mean(count)) 

#This was!
ggplot()+
 geom_line(data = attorney_quarter_sum, aes(x = quarter , y = average, color = Group),
           linewidth = 1)+
  geom_point(data = attorney_quarter_sum, aes(x = quarter , y = average, color = Group))+
  scale_x_yearqtr(format = "%YQ%q", n = 15)+
  scale_color_manual(values = c(
    "Trial Division" = "navy",
    "No Units Entered" = "lightblue",
    "Major Crimes" = "tan4",
    "Special Units" = "blue"
  ))+
  labs(x = "Time when case was received",
       y = "Count",
       title = "Quarterly Average Caseload per Attorney Across Different Divisions (Case-Level)")+
  geom_rect(data= covid_period, aes(xmin = x1, xmax = x2, ymin = y1, ymax = y2), fill = "orange",
            alpha = 0.3)+
  theme_bw() + 
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.text = element_text(size = 7),   
    legend.title = element_text(size = 8),
    legend.position = "bottom"              
  )
```

###### AI use: 
prompt: I have two datasets: one df has a column of code (1-1-1(a), for example), and the other df has reference code (1-1-1). i want to filter for all rows in the df that contain the mother-level code in the refernece df
response: 
```{r}
pattern <- ref_df$ref_code |>
  sapply(FUN = \(x) gsub("([.\\-])", "\\\\\\1", x)) |>
  paste(collapse = "|")

filtered_df <- df |>
  filter(grepl(pattern, code))
```

```{r}
pattern <- svf_code$offense_code |>
  sapply(FUN = \(x) gsub("([.\\-])", "\\\\\\1", x)) |>
  paste(collapse = "|")

post_indictment_svf = post_indictment_cr |>
  filter(grepl(pattern, offense_code),
         Unit != "Trial Division",
         !Unit %in% special_labels) #549
```




```{r}
# assign colors to each division
custom_colors <- c(
  "D1 odd"  = "#00008B",
  "D1 even" = "#00008B",

  "D2 odd"  = "#8B4513",
  "D2 even" = "#8B4513",

  "D3 odd"  = "#FF8C00",
  "D3 even" = "#FF8C00",

  "D4 odd"  = "#2F4F4F",
  "D4 even" = "#2F4F4F",

  "D5 odd"  = "#006400",
  "D5 even" = "#006400",

  "D6 odd"  = "#8B0000",
  "D6 even" = "#8B0000",

  "D7 odd"  = "#4B0082",
  "D7 even" = "#4B0082",

  "D8 odd"  = "#00688B",
  "D8 even" = "#00688B",

  "D9 odd"  = "#556B2F",
  "D9 even" = "#556B2F",

  "D10 odd"  = "#8B008B",
  "D10 even" = "#8B008B"
)


```



```{r}
# define function for plotting graphs (specific breakdown, by each quarter)
filter_by_division <- function(df) {
  df <- df %>% 
    mutate(quarter = as.yearqtr(date_assigned,  format = "%Y-%m-%d"))
quarters <- unique(df$quarter)

   for (i in 1:length(quarters)) {
   q <- quarters[i]
  
  filtered <- df %>%
    filter(quarter == q) %>%
            group_by(Attorney) %>%
            summarise(n = n(), .groups = "drop") %>%
            arrange(Attorney)

        name <- paste(q, sep = "_")

        p <- ggplot(filtered, aes(x = Attorney, y = n, fill = Attorney)) +
            geom_col() +
            scale_fill_manual(values = custom_colors) +
            labs(
                title = paste("Initial Attorney Assignment -", q),
                x = "Division",
                y = "Count"
            ) +
          ylim(0, 75)+
            theme_bw(base_size = 10) +
            coord_cartesian(ylim = c(0, 75)) +
            theme(
                legend.position = "none",
                axis.text.x = element_text(angle = 20, hjust = 1)
            )
      ggsave(filename = paste0("Trial_division_",  q, ".png"),
      plot = p,
      width = 8, height = 5)
      print(p)
    }
}
```
 
```{r} 
#Graphing
filter_by_division(post_indictment_trial)

```

```{r} 
# define function for plotting graphs (specific breakdown, by each quarter)
filter_by_unit <- function(df) {
  df <- df %>% 
    mutate(quarter = as.yearqtr(date_assigned,  format = "%Y-%m-%d"))
quarters <- unique(df$quarter)

      for (i in 1:length(quarters)) {
   q <- quarters[i]

        filtered <- df %>%
            filter(quarter == q) %>%
            group_by(Attorney, Unit) %>%
            summarise(count = n(), .groups = "drop") %>%
            arrange(Unit)

        name <- paste(q, sep = "_")

        p <- ggplot(filtered, aes(x = Attorney, y = count, fill = Unit)) +
            geom_col() +
           scale_fill_manual(values = c("DVSA" = "wheat", 
                                "CCU" = "#daa520",
                                "CAEPAD" = "tan4", 
                                "HTIC" = "slategrey" ,
                                "ACU" = "cornflowerblue",
                                "FVPU" = "#4169B0",
                                "Homicide/Gangs" = "orange"))+
          ylim(0, 75)+
            labs(
                title = paste("Special Units Caseload Distribution -", q),
                x = "Unit",
                y = "Count"
            ) +
          
            theme_bw(base_size = 10) +
            theme(
                legend.position = "right",
                axis.text.x = element_text(angle = 20, hjust = 1)
            )
      ggsave(filename = paste0("special_unit_",  q, ".png"),
      plot = p,
      width = 8, height = 5)
      print(p)
    }
}

filter_by_unit(post_indictment_special_units)
post_indictment_special_units %>% filter(Unit == "FVPU")
```

