---
title: "DA_SCR_Scheduling_Ryan_0430_VFinal"
author: "Ryan Wang"
date: "2026-04-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load necessary packages
library(tidyverse)
library(janitor)
library(ggplot2)
library(stringr)
library(lubridate)
getwd()
# Load data
hearing_data <- read.csv("/Users/karizhou/Desktop/SPRING 26/Capstone/Data/Hearing Data - Comments(Hearing Data - Comments).csv")

cases_charged_data <- read.csv("/Users/karizhou/Desktop/SPRING 26/Capstone/Data/Cases Charged Data.csv")
case_assignment_data <- read.csv("/Users/karizhou/Desktop/SPRING 26/Capstone/Data/Case Division Data _ Post Standing Order _ Post-Indictment Assignments(Sheet1).csv")
bench_warrant_data <- read.csv("/Users/karizhou/Desktop/SPRING 26/Capstone/Data/Bench Warrant Data(Bench Warrant Data).csv")
cases_closed_data <- read.csv("/Users/karizhou/Desktop/SPRING 26/Capstone/Data/Cases Closed Data.csv")
```

# Data Exploration & Cleaning

## Renaming Variables

```{r}
# Rename column names to be in snake_case
hearing_data <- clean_names(hearing_data)
cases_charged_data <- clean_names(cases_charged_data)
case_assignment_data <- clean_names(case_assignment_data)
bench_warrant_data <- clean_names(bench_warrant_data)
cases_closed_data <- clean_names(cases_closed_data)
```

## Cleaning Dates

```{r}
# Examine variables
str(hearing_data)
str(cases_charged_data)
str(case_assignment_data)
str(bench_warrant_data)
str(cases_closed_data)

# Change date variables to become dates
hearing_data$hearing_date <- as.Date(hearing_data$hearing_date, format = "%m/%d/%Y")

cases_charged_data$offense_date <- as.Date(cases_charged_data$offense_date, format = "%m/%d/%y")
cases_charged_data$arrest_date <- as.Date(cases_charged_data$offense_date, format = "%m/%d/%y")
cases_charged_data$received_date <- as.Date(cases_charged_data$received_date, format = "%m/%d/%y")
cases_charged_data$event_date <- as.Date(cases_charged_data$event_date, format = "%m/%d/%y")

case_assignment_data$division_assignment_time_stamp_created <- as.Date(case_assignment_data$division_assignment_time_stamp_created, format = "%m/%d/%Y")
case_assignment_data$division_assignment_time_stamp_updated <- as.Date(case_assignment_data$division_assignment_time_stamp_updated, format = "%m/%d/%Y")

bench_warrant_data$event_date <- as.Date(bench_warrant_data$event_date, format = "%m/%d/%Y")

cases_closed_data$offense_date <- as.Date(cases_closed_data$offense_date, format = "%m/%d/%y")
cases_closed_data$arrest_date <- as.Date(cases_closed_data$arrest_date, format = "%m/%d/%y")
cases_closed_data$received_date <- as.Date(cases_closed_data$received_date, format = "%m/%d/%y")
cases_closed_data$event_date <- as.Date(cases_closed_data$event_date, format = "%m/%d/%y")

# Verify cleaning
str(hearing_data)
str(cases_charged_data)
str(case_assignment_data)
str(bench_warrant_data)
str(cases_closed_data)
```

## Filtering for Trial Division
```{r}
# Only include cases listed as trial division
cases_charged_trial_division <- cases_charged_data %>%
  filter(cases_charged_data$unit == "Trial Division")

cases_closed_trial_division <- cases_closed_data %>%
  filter(cases_closed_data$unit == "Trial Division")

# Remove original data to tidy up environment
rm(cases_charged_data)
rm(cases_closed_data)
```

## Checking for Duplicates
```{r}
# Check for duplicates and view duplicates
sum(duplicated(hearing_data))

sum(duplicated(cases_charged_trial_division))
cases_charged_trial_division[duplicated(cases_charged_trial_division), ]

sum(duplicated(case_assignment_data))
case_assignment_data[duplicated(case_assignment_data), ]

sum(duplicated(bench_warrant_data))

sum(duplicated(cases_closed_trial_division))
cases_closed_trial_division[duplicated(cases_closed_trial_division), ]

# Remove duplicates
cases_charged_trial_division <- cases_charged_trial_division[!duplicated(cases_charged_trial_division), ]

case_assignment_data <- case_assignment_data[!duplicated(case_assignment_data), ]

cases_closed_trial_division <- cases_closed_trial_division[!duplicated(cases_closed_trial_division), ]

# Verify cleaning
sum(duplicated(cases_charged_trial_division))
sum(duplicated(case_assignment_data))
sum(duplicated(cases_closed_trial_division))
```

## Checking for Errors
```{r}
# Inspect data
summary(hearing_data)
summary(cases_charged_trial_division)
summary(case_assignment_data)
summary(bench_warrant_data)
summary(cases_closed_trial_division)

# Filter for indicted cases
indicted_cases_trial_division <- cases_charged_trial_division %>%
  filter(cases_charged_trial_division$event == "Indictment")

# See if cases were indicted before they were received
indicted_cases_trial_division[indicted_cases_trial_division$event_date < indicted_cases_trial_division$received_date, ]

# See if cases were closed before they were received
cases_closed_trial_division[cases_closed_trial_division$event_date < cases_closed_trial_division$received_date, ]

# Remove errors
indicted_cases_trial_division <- indicted_cases_trial_division[indicted_cases_trial_division$event_date >= indicted_cases_trial_division$received_date | is.na(indicted_cases_trial_division$event_date), ]

cases_closed_trial_division <- cases_closed_trial_division[cases_closed_trial_division$event_date >= cases_closed_trial_division$received_date | is.na(cases_closed_trial_division$event_date), ]

# Verify cleaning
indicted_cases_trial_division[indicted_cases_trial_division$event_date < indicted_cases_trial_division$received_date, ]
cases_closed_trial_division[cases_closed_trial_division$event_date < cases_closed_trial_division$received_date, ]

# Remove data to tidy up environment
rm(cases_charged_trial_division)
```

## Case to Case-Defendant Level
```{r}
# Turn indicted cases into case-defendant level (summarize charges)
indicted_cases_trial_division_defendant_level <- indicted_cases_trial_division %>%
  group_by(case_number, party_id) %>%
  summarise(
    defendant = unique(defendant),
    indictment_summary = paste(event_date, collapse = "; "), # Summarize indictment dates, separating with semicolons
    charges_summary = paste(offense, collapse = "; "), # Summarize charges, separating with semicolons
    charges_code_summary = paste(offense_code, collapse = "; "), # Summarize charge codes, separating with semicolons
    .groups = "drop"
  )

# Remove data to tidy up environment
rm(indicted_cases_trial_division)
```

```{r}
# Turn closed cases into case-defendant level (summarize charges)
cases_closed_trial_division_defendant_level <- cases_closed_trial_division %>%
  group_by(case_number, party_id) %>%
  summarise(
    defendant = unique(defendant),
    closure_summary = paste(event_date, collapse = "; "), # Summarize closure dates, separating with semicolons
    charges_summary = paste(offense, collapse = "; "), # Summarize charges, separating with semicolons
    charges_code_summary = paste(offense_code, collapse = "; "), # Summarize charge codes, separating with semicolons
    .groups = "drop"
  )

# Remove data to tidy up environment
rm(cases_closed_trial_division)
```

```{r}
# Turn case assignments into case-defendant level (summarize assignments)
case_assignment_data_defendant_level <- case_assignment_data %>%
  mutate(assignment_time = division_assignment_time_stamp_created) %>%
  group_by(case_number, defendant_name, party_id) %>%
  slice_max(order_by = assignment_time, n = 1, with_ties = FALSE) %>% # Use most recent assignment time
  ungroup()

# Remove data to tidy up environment
rm(case_assignment_data)
```

```{r}
# Turn bench warrant data into case-defendant level (summarize bench warrants)
bench_warrant_data_defendant_level <- bench_warrant_data %>%
  group_by(case_number, defendant, party_id) %>%
  summarise(
    num_bench_warrants = n(),
    .groups = "drop"
  )

# Remove data to tidy up environment
rm(bench_warrant_data)
```

## Joining Datasets

```{r}
# Create master dataset
master_data <- hearing_data %>%
  left_join(indicted_cases_trial_division_defendant_level, by = c("case_number", "party_id")) %>%
  left_join(cases_closed_trial_division_defendant_level, by = c("case_number", "party_id")) %>%
  left_join(case_assignment_data_defendant_level, by = c("case_number", "party_id")) %>%
  left_join(bench_warrant_data_defendant_level, by = c("case_number", "party_id"))
```

# Duration from Indictment to Arraignment

## Removing Bench Warrants

```{r}
# Filter master data to exclude bench warrants
master_data <- master_data %>%
  filter(is.na(num_bench_warrants))
```

## Descriptive Statistics and Distribution of Duration from Indictment to Arraignment

```{r}
# Create helper function for converting days to years + months
days_to_years_months <- function(days) {
  years <- floor(days / 365) # Divide days by 365, rounding down to the nearest whole number
  months <- floor((days %% 365) / 30) # Convert the remainder to months, again rounding down
  
  case_when(
    years == 0 & months == 0 ~ paste("< 1 mo"), # Less than 1 month
    years == 0 ~ paste(months, "mo"), # Only months
    months == 0 ~ paste(years, "yr"), # Only years
    TRUE ~ paste0(years, "yr ", months, "mo") # Years and months
  )
}

# Create helper function for converting days to months
days_to_months <- function(days) {
  months <- round(days / 30) # Convert days to months
  ifelse(months == 1, "1 month", paste0(months, " months")) # Create human-readable text
}
```

```{r}
# Filter master data to exclude data without a creation date
master_data <- master_data %>%
  filter(!is.na(division_assignment_time_stamp_created)) %>%
  filter(division_assignment_time_stamp_created >= as.Date("2022-08-01")) # Make sure the data is post-standing order

# Make sure the data is a proper date object
master_data$hearing_date <- as.Date(master_data$hearing_date)

# Create variable for number of charges
master_data$n_charges <- str_count(master_data$charges_summary.x, ";") + 1

# Filter for only arraignment hearings
master_data_arraignments <- master_data %>%
  filter(hearing_type == "Arraignment") %>%
  group_by(case_number, party_id) %>%
  slice_max(order_by = hearing_date, n = 1, with_ties = FALSE) %>% # Keep only the row with the latest hearing_date
  ungroup()

# Create variable for EARLIEST indictment
# Split the strings into pieces
master_data_arraignments$earliest_indictment <- sapply(strsplit(master_data_arraignments$indictment_summary, ";") # Loop over each row
                                                       , function(x) {
  dates <- as.Date(trimws(x), format = "%Y-%m-%d") # Clean and convert to date format
  if(all(is.na(dates))) return(NA) # Handle rows with no valid dates
  min(dates, na.rm = TRUE) # Take the earliest date
})

# Convert back to date
master_data_arraignments$earliest_indictment <- 
  as.Date(master_data_arraignments$earliest_indictment, origin = "1970-01-01")

# Calculate duration between indictment and arraignment
master_data_arraignments$indictment_arraignment_duration <- as.numeric(
  as.Date(master_data_arraignments$hearing_date) - master_data_arraignments$earliest_indictment
)
```

```{r}
summary(master_data_arraignments$indictment_arraignment_duration)
```

```{r}
master_data_arraignments <- master_data_arraignments %>% # Create bins
  mutate(duration_bin = case_when(
    indictment_arraignment_duration < 30  ~ "< 1 mo",
    indictment_arraignment_duration < 90  ~ "1–3 mo",
    indictment_arraignment_duration < 180 ~ "3–6 mo",
    indictment_arraignment_duration < 365 ~ "6–12 mo",
    indictment_arraignment_duration >= 365 ~ "1+ years",
    TRUE ~ NA_character_ # If nothing above matched, return an NA
  ))

# Set the order explicitly (otherwise bins will be alphabetized)
master_data_arraignments$duration_bin <- factor(
  master_data_arraignments$duration_bin,
  levels = c("< 1 mo", "1–3 mo", "3–6 mo", "6–12 mo", "1+ years")
)
```

```{r}
# Plot as a categorical histogram/bar chart
master_data_arraignments %>%
  filter(!is.na(duration_bin)) %>%
  
  ggplot(aes(x = duration_bin)) +
  geom_bar(fill = "steelblue", color = "white") +
  
  labs(
    x = "Time from Indictment to Arraignment",
    y = "Number of Cases",
    title = "Distribution of Indictment to Arraignment Duration",
    subtitle = "Case-Defendant Level, Trial Division, No Bench Warrants, Post-Standing Order"
  ) +
  
  theme_minimal()
```

# Number of Charges and Duration from Indictment to Arraignment

```{r}
# Table
n_charges_duration_table <- master_data_arraignments %>%
  group_by(n_charges) %>%
  summarise(
    mean_duration = mean(indictment_arraignment_duration, na.rm = TRUE),
    median_duration = median(indictment_arraignment_duration, na.rm = TRUE),
    n = n()
  )
```

```{r}
# Combine 20+ charges
master_data_arraignments %>%
  mutate(charge_category = case_when(
    n_charges >= 20 ~ 20, # If a case has 20 or more charges, it is assigned 20
    TRUE ~ n_charges # Otherwise it keeps the original number of charges
  )) %>%
  group_by(charge_category) %>%
  summarise(
    n_cases = n(),
    mean_indictment_arraignment_duration = mean(indictment_arraignment_duration, na.rm = TRUE)
  ) %>%
  filter(n_cases >= 10) %>% # Filter for charge counts that are 10 or more
  
  # Graphing
  ggplot(aes(x = charge_category, y = mean_indictment_arraignment_duration)) +
  geom_line(color = "darkblue") +
  geom_point() +
  scale_x_continuous(
    breaks = c(1:19, 20),
    labels = c(1:19, "20+")
  ) +
  scale_y_continuous(
    breaks = seq(0, max(master_data_arraignments$indictment_arraignment_duration, na.rm = TRUE), by = 30),
    labels = days_to_months
  ) +
  labs(
    x = "Number of Charges",
    y = "Average Duration",
    title = "Indictment to Arraignment Duration by Number of Charges",
    subtitle = "Case-Defendant Level, Trial Division, No Bench Warrants, Post-Standing Order"
  ) +
  theme_minimal()
```
```{r}
# Boxplot version
master_data_arraignments %>%
  filter(!is.na(n_charges), indictment_arraignment_duration >= 0) %>%  # Filter out NAs and negative durations
  mutate(num_charges_display = ifelse(n_charges >= 11, 11, n_charges)) %>% # Combine 11+ charges
  group_by(num_charges_display) %>%
  filter(n() >= 10) %>% # Only include charge counts with 10 or more cases
  ungroup() %>%
  
  ggplot(aes(x = factor(num_charges_display), y = indictment_arraignment_duration)) +
  geom_boxplot() +
  scale_x_discrete(labels = c(1:10, "11+")) +
  scale_y_continuous(
  breaks = c(
    0,
    180,
    365,
    545,
    730,
    910,
    1095,
    1275,
    1460
  ),
  labels = c(
    "<1 mo",
    "6 mo",
    "1 yr",
    "1 yr 6 mo",
    "2 yr",
    "2 yr 6 mo",
    "3 yr",
    "3 yr 6 mo",
    "4 yr"
  )
) +
  labs(x = "Number of Charges",
       y = "Duration",
       title = "Indictment to Arraignment Duration by Number of Charges",
       subtitle = "Case-Defendant Level, Trial Division, No Bench Warrants, Post-Standing Order") +
  theme_minimal()
```
```{r}
# Remove data to tidy environment
rm(n_charges_duration_table)
```

# Visualizing the Relationship Between the Number of Defendants and Indictment-Arraignment Duration

```{r}
# Create variable for number of defendants
master_data_arraignments <- master_data_arraignments %>%
  group_by(case_number) %>%
  mutate(n_defendants = n()) %>%
  ungroup()
```

```{r}
# Table
n_defendants_duration_table <- master_data_arraignments %>%
  group_by(n_defendants) %>%
  summarise(
    mean_indictment_araignment_duration = mean(indictment_arraignment_duration, na.rm = TRUE),
    median_indictment_arraignment_duration = median(indictment_arraignment_duration, na.rm = TRUE),
    n = n()
  )
```

```{r}
# Combine 20+ defendants
master_data_arraignments %>%
  mutate(defendant_category = case_when(
    n_defendants >= 20 ~ 20, # If a case has 20 or more defendants, it is assigned 20
    TRUE ~ n_defendants # Otherwise it keeps the original number of defendants
  )) %>%
  group_by(defendant_category) %>%
  summarise(
    n_cases = n(),
    mean_indictment_arraignment_duration = mean(indictment_arraignment_duration, na.rm = TRUE)
  ) %>%
  # Filter for defendant counts 10 or more
  filter(n_cases >= 10) %>%
  
  # Graphing
  ggplot(aes(x = defendant_category, y = mean_indictment_arraignment_duration)) +
  geom_line(color = "darkblue") +
  geom_point() +
  scale_x_continuous(
    breaks = c(1:19, 20),
    labels = c(1:19, "20+")
  ) +
  scale_y_continuous(
    breaks = seq(0, max(master_data_arraignments$indictment_arraignment_duration, na.rm = TRUE), by = 30),
    labels = days_to_years_months
  ) +
  labs(
    x = "Number of Defendants",
    y = "Average Duration",
    title = "Indictment to Arraignment Duration by Number of Defendants",
    subtitle = "Case-Defendant Level, Trial Division, No Bench Warrants, Post-Standing Order"
  ) +
  theme_minimal()
```

```{r}
# Boxplot version
master_data_arraignments %>%
  filter(!is.na(n_defendants), # Exclude NA's
         indictment_arraignment_duration >= 0) %>% # Exclude negative durations
  
  # Collapse 11+ defendants
  mutate(num_defendants_display = ifelse(n_defendants >= 11, 11, n_defendants)) %>%
  
  # Compute group sizes BEFORE filtering anything out
  group_by(num_defendants_display) %>%
  mutate(group_n = n()) %>%
  ungroup() %>%
  
  # Keep only groups with at least 10 cases
  filter(group_n >= 10) %>%
  
  ggplot(aes(x = factor(num_defendants_display),
             y = indictment_arraignment_duration)) +
  geom_boxplot() +
  
  scale_x_discrete(
    labels = function(x) ifelse(x == "11", "11+", x)
  ) +
  
  scale_y_continuous(
    breaks = c(0, 180, 365, 545, 730, 910, 1095, 1275, 1460),
    labels = c("<1 mo", "6 mo", "1 yr", "1 yr 6 mo",
               "2 yr", "2 yr 6 mo", "3 yr", "3 yr 6 mo", "4 yr")
  ) +
  
  labs(
    x = "Number of Defendants",
    y = "Duration",
    title = "Indictment to Arraignment Duration by Number of Defendants",
    subtitle = "Case-Defendant Level, Trial Division, No Bench Warrants, Post-Standing Order"
  ) +
  
  theme_minimal()
```

```{r}
# Remove data to tidy environment
rm(n_defendants_duration_table)
```

# Visualizing Indictment-Arraignment Durations by Division

```{r}
# Include only Divisions 1-10, filter out NA's, remove negative durations
master_data_arraignments <- master_data_arraignments %>%
  filter(!division %in% c("Superior Court", "Superior Pre-File")) %>%
  filter(!is.na(indictment_arraignment_duration)) %>%
  filter(indictment_arraignment_duration >= 0)

# Order the Divisions from 1 to 10
master_data_arraignments$division <- factor(
  master_data_arraignments$division,
  levels = unique(
    master_data_arraignments$division[
      order(as.numeric(gsub("[^0-9]", "", master_data_arraignments$division)))
    ]
  )
)
```

```{r}
# Boxplot graph
ggplot(master_data_arraignments, 
       aes(x = division, 
           y = indictment_arraignment_duration, 
           fill = division)) +

  geom_boxplot(outlier.shape = 21, outlier.size = 2, outlier.alpha = 0.6) +

  scale_fill_manual(values = c(
    "cornflowerblue", "cadetblue1", "forestgreen", "yellowgreen", "darkolivegreen",
    "tan4", "wheat4", "firebrick4","orange", "moccasin"
  ), guide = "none") +

  scale_y_continuous(
  breaks = seq(0, 4 * 365, by = 182.5),  # every ~6 months
  labels = days_to_years_months
) +

  labs(
    title = "Indictment to Arraignment Duration by Division",
    subtitle = "Case-Defendant Level, Trial Division, No Bench Warrants, Post-SO",
    x     = "Division",
    y     = "Indictment to Arraignment Duration"
  ) +

  theme_minimal(base_size = 13) +
  theme(
    plot.title    = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, color = "grey50"),
    legend.position = "bottom",
    panel.grid.major.x = element_blank(),
    axis.text.x        = element_text(angle = 30, hjust = 1, size = 9)
  )
```

```{r}
# Basic ANOVA test
anova_model_indictment_arraignment <- aov(indictment_arraignment_duration ~ as.factor(division), 
                   data = master_data_arraignments)

# View ANOVA results
summary(anova_model_indictment_arraignment)
```

```{r}
# Post-hoc testing (Tukey's Honest Significant Difference test)
tukey_results_indictment_arraignment <- TukeyHSD(anova_model_indictment_arraignment)
print(tukey_results_indictment_arraignment)

# Extract results
tukey_df_indictment_arraignment <- as.data.frame(tukey_results_indictment_arraignment$`as.factor(division)`)

# Filter for significant differences
significant_pairs_indictment_arraignment <- tukey_df_indictment_arraignment %>%
  filter(`p adj` < 0.05) %>%
  arrange(`p adj`)

print(significant_pairs_indictment_arraignment)
```
```{r}
# Remove data to tidy environment
rm(anova_model_indictment_arraignment)
rm(significant_pairs_indictment_arraignment)
rm(tukey_df_indictment_arraignment)
rm(tukey_results_indictment_arraignment)
```

# Time Between Hearings

```{r}
# Filter out bond hearings and convert hearing_date to date variable
master_data <- master_data %>%
  filter(!hearing_type %in% c("Bond Hearings", "Bond Reduction Hearing")) %>%
  mutate(hearing_date = as.Date(hearing_date))

# Select the EARLIEST arraignment
earliest_arraignments <- master_data %>%
  filter(hearing_type == "Arraignment") %>%
  group_by(case_number, party_id) %>%
  slice_min(order_by = hearing_date, n = 1, with_ties = FALSE) %>%
  ungroup()

# Find the next hearing after arraignment
arraignment_to_next <- earliest_arraignments %>%
  left_join(
    master_data,
    by = c("case_number", "defendant.x", "party_id"),
    suffix = c("_arraignment", "_next")
  ) %>%
  filter(hearing_date_next > hearing_date_arraignment) %>% # Keep only hearings AFTER arraignment
  group_by(case_number, party_id) %>% # Pick the NEXT immediate hearing
  slice_min(order_by = hearing_date_next, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  # Compute time difference
  mutate(
    days_to_next_hearing = as.numeric(
      hearing_date_next - hearing_date_arraignment
    )
  )
```

```{r}
# Find the mean and median
summary(arraignment_to_next$days_to_next_hearing)
```

```{r}
# Join dataset with division numbers and filter out bench warrants
arraignment_to_next_divisions_no_bench_warrants <- arraignment_to_next %>%
  left_join(bench_warrant_data_defendant_level, by = c("case_number", "party_id")) %>%
  left_join(case_assignment_data_defendant_level, by = c("case_number", "party_id")) %>%
  filter(is.na(num_bench_warrants))
```

```{r}
# Filter for only Divisions 1-10, exclude NA's and negative durations
arraignment_to_next_divisions_no_bench_warrants <- 
  arraignment_to_next_divisions_no_bench_warrants %>%
  filter(!division %in% c("Superior Court", "Superior Pre-File")) %>%
  filter(!is.na(days_to_next_hearing)) %>%
  filter(days_to_next_hearing >= 0)

# Factor the division variable to properly sort from 1-10
arraignment_to_next_divisions_no_bench_warrants$division <- factor(
  arraignment_to_next_divisions_no_bench_warrants$division,
  levels = unique(
    arraignment_to_next_divisions_no_bench_warrants$division[
      order(as.numeric(gsub("[^0-9]", "", 
            arraignment_to_next_divisions_no_bench_warrants$division)))
    ]
  )
)
```

```{r}
# Boxplot
ggplot(arraignment_to_next_divisions_no_bench_warrants, 
       aes(x = division, 
           y = days_to_next_hearing, 
           fill = division)) +

  geom_boxplot(outlier.shape = 21, outlier.size = 2, outlier.alpha = 0.6) +

  scale_fill_manual(values = c(
    "cornflowerblue", "cadetblue1", "forestgreen", "yellowgreen", "darkolivegreen",
    "tan4", "wheat4", "firebrick4","orange", "moccasin"
  ), guide = "none") +

  scale_y_continuous(
    breaks = seq(0, 3 * 365, by = 182.5),
    labels = days_to_years_months
  ) +

  labs(
    title = "Arraignment to Subsequent Hearing Duration by Division",
    subtitle = "Case-Defendant Level, Trial Division, No Bench Warrants, Post-SO",
    x = "Division",
    y = "Days from Arraignment to Subsequent Hearing"
  ) +

  theme_minimal(base_size = 13) +
  theme(
    plot.title    = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, color = "grey50"),
    legend.position = "bottom",
    panel.grid.major.x = element_blank(),
    axis.text.x        = element_text(angle = 30, hjust = 1, size = 9)
  )
```

```{r}
# Basic ANOVA test
anova_model_arraignment_next_hearing <- aov(days_to_next_hearing ~ as.factor(division), 
                   data = arraignment_to_next_divisions_no_bench_warrants)

# View ANOVA results
summary(anova_model_arraignment_next_hearing)
```

```{r}
# Post-hoc testing (Tukey's Honest Significant Difference test)
tukey_results_arraignment_next_hearing <- TukeyHSD(anova_model_arraignment_next_hearing)
print(tukey_results_arraignment_next_hearing)

# Extract results
tukey_df_arraignment_next_hearing <- as.data.frame(tukey_results_arraignment_next_hearing$`as.factor(division)`)

# Filter for significant differences
significant_pairs_arraignment_next_hearing <- tukey_df_arraignment_next_hearing %>%
  filter(`p adj` < 0.05) %>%
  arrange(`p adj`)

print(significant_pairs_arraignment_next_hearing)
```
```{r}
# Remove data to tidy environment
rm(anova_model_arraignment_next_hearing)
rm(arraignment_to_next)
rm(arraignment_to_next_divisions_no_bench_warrants)
rm(earliest_arraignments)
rm(significant_pairs_arraignment_next_hearing)
rm(tukey_df_arraignment_next_hearing)
rm(tukey_results_arraignment_next_hearing)
```

# Case Closure Times

```{r}
# Split the closure summary into individual closure dates
master_data <- master_data %>%
  mutate(
    latest_closure_date = str_split(closure_summary, ";\\s*") %>% # Split into pieces at semicolons
      map(~ as.Date(.x, format = "%Y-%m-%d")) %>% # Convert strings to Date objects
      # Handle missing/empty lists
      map(~ {
        if (all(is.na(.x)) || length(.x) == 0) {
          return(NA)
        }
        max(.x, na.rm = TRUE)
      }) %>%
      # Convert into clean Date
      map_chr(~ as.character(.x)) %>%
      as.Date()
  )

# Calculate duration from assignment to latest closure
master_data <- master_data %>%
  mutate(
    assignment_to_closure_duration = as.numeric(difftime(latest_closure_date, 
                                                          assignment_time, 
                                                          units = "days"))
  )

# Check the results
master_data %>%
  select(closure_summary, latest_closure_date, assignment_time, assignment_to_closure_duration) %>%
  head(10)
```

```{r}
# How many NAs and negatives?
master_data %>%
  summarise(
    total_cases = n(),
    na_closure_summary = sum(is.na(closure_summary)),
    na_latest_closure = sum(is.na(latest_closure_date)),
    na_assignment = sum(is.na(assignment_time)),
    negative_duration = sum(assignment_to_closure_duration < 0, na.rm = TRUE),
    positive_duration = sum(assignment_to_closure_duration >= 0, na.rm = TRUE)
  )

# Look at some examples of negative durations
master_data %>%
  filter(assignment_to_closure_duration < 0) %>%
  select(assignment_time, closure_summary, latest_closure_date, assignment_to_closure_duration) %>%
  head(20)

# Look at some NAs
master_data %>%
  filter(is.na(assignment_to_closure_duration)) %>%
  select(assignment_time, closure_summary, latest_closure_date, assignment_to_closure_duration) %>%
  head(20)
```

```{r}
# Only keep cases where assignment came before closure
master_data <- master_data %>%
  filter(!is.na(assignment_time), 
         !is.na(latest_closure_date),
         assignment_to_closure_duration >= 0)

# Check how many cases remain
nrow(master_data)

# Descriptive statistics
summary(master_data$assignment_to_closure_duration)
```

# Case Closure Times by Division
```{r}
# Include only Divisions 1-10, filter for CR cases, remove NA's and negative durations
master_data <- master_data %>%
  filter(!division %in% c("Superior Court", "Superior Pre-File")) %>%
  filter(grepl("CR", case_number, ignore.case = TRUE)) %>%
  filter(!is.na(assignment_to_closure_duration)) %>%
  filter(assignment_to_closure_duration >= 0)

# Order the Divisions from 1 to 10
master_data$division <- factor(
  master_data$division,
  levels = unique(
    master_data$division[
      order(as.numeric(gsub("[^0-9]", "", master_data$division)))
    ]
  )
)
```


```{r}
# Boxplot graph
ggplot(master_data, 
       aes(x = division, 
           y = assignment_to_closure_duration, 
           fill = division)) +

  geom_boxplot(outlier.shape = 21, outlier.size = 2, outlier.alpha = 0.6) +

  scale_fill_manual(values = c(
    "cornflowerblue", "cadetblue1", "forestgreen", "yellowgreen", "darkolivegreen",
    "tan4", "wheat4", "firebrick4","orange", "moccasin"
  ), guide = "none") +

  scale_y_continuous(
  breaks = seq(0, 3 * 365, by = 182.5),  # every ~6 months
  labels = days_to_years_months
) +

  labs(
    title = "Assignment to Closure Duration by Division",
    subtitle = "Case-Defendant Level, Trial Division CR Cases, No Bench Warrants, Post-SO",
    x     = "Division",
    y     = "Assignment to Closure Duration"
  ) +

  theme_minimal(base_size = 13) +
  theme(
    plot.title    = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, color = "grey50"),
    legend.position = "bottom",
    panel.grid.major.x = element_blank(),
    axis.text.x        = element_text(angle = 30, hjust = 1, size = 9)
  )
```

```{r}
# Basic ANOVA test
anova_model_assignment_closure <- aov(assignment_to_closure_duration ~ as.factor(division), 
                   data = master_data)

# View ANOVA results
summary(anova_model_assignment_closure)
```

```{r}
# Post-hoc testing (Tukey's Honest Significant Difference test)
tukey_results_assignment_closure <- TukeyHSD(anova_model_assignment_closure)
print(tukey_results_assignment_closure)

# Extract results
tukey_df_assignment_closure <- as.data.frame(tukey_results_assignment_closure$`as.factor(division)`)

# Filter for significant differences
significant_pairs_assignment_closure <- tukey_df_assignment_closure %>%
  filter(`p adj` < 0.05) %>%
  arrange(`p adj`)

print(significant_pairs_assignment_closure)
```

```{r}
# Remove data to tidy environment
rm(anova_model_assignment_closure)
rm(significant_pairs_assignment_closure)
rm(tukey_df_assignment_closure)
rm(tukey_results_assignment_closure)
```

# Hearings Until Case Closure

```{r}
# Count non-bond hearings per case (CR cases only)
hearings_per_case <- master_data %>%
  filter(!hearing_type %in% c("Bond Hearing", "Bond Reduction Hearing")) %>%
  filter(grepl("CR", case_number, ignore.case = TRUE)) %>%  # Only CR cases
  group_by(case_number) %>%
  summarise(
    n_hearings = n(),
    hearing_types = paste(unique(hearing_type), collapse = "; ")
  )

# Summary statistics
summary(hearings_per_case$n_hearings)
table(hearings_per_case$n_hearings)
```

```{r}
# Create a binned version of n_hearings
hearings_per_case_binned <- hearings_per_case %>%
  mutate(
    n_hearings_binned = ifelse(n_hearings >= 20, "20+", as.character(n_hearings)),
    n_hearings_binned = factor(n_hearings_binned, 
                               levels = c(as.character(1:19), "20+"))
  )

# Histogram with 20+ combined
ggplot(hearings_per_case_binned, aes(x = n_hearings_binned)) +
  geom_bar(fill = "darkblue", alpha = 0.7) +
  labs(
    title = "Distribution of Hearings Among Closed CR Cases",
    subtitle = "Case Level, Trial Division, Post-Standing Order, Excludes Bond Hearings",
    x = "Number of Hearings",
    y = "Number of Cases"
  ) +
  theme_minimal()
```

# Hearing Until Case Closure by Division
```{r}
# Calculate median hearings per case for each division
hearings_by_division <- master_data %>%
  # Filter out bond hearings
  filter(!hearing_type %in% c("Bond Hearing", "Bond Reduction Hearing")) %>%
  # Filter for CR cases
  filter(grepl("CR", case_number, ignore.case = TRUE)) %>%
  # Count hearings per case
  group_by(case_number, division) %>%
  summarise(n_hearings = n(), .groups = "drop") %>%
  # Remove NA divisions and fake divisions
  filter(!is.na(division),
         !division %in% c("Superior Pre-File", "Superior Court")) %>%
  # Calculate median by division
  group_by(division) %>%
  summarise(
    n_cases = n(),
    median_hearings = median(n_hearings, na.rm = TRUE)
  ) %>%
  # Order divisions 1-10 (with "Division " prefix)
  mutate(division = factor(division, 
                          levels = paste("Division", 1:10))) %>%
  arrange(division)

# View the results
print(hearings_by_division)
```

```{r}
ggplot(hearings_by_division, aes(x = division, y = median_hearings)) +
  geom_col(fill = "steelblue") +
  labs(
    title = "Median Hearings Among Closed CR Cases by Division",
    subtitle = "Case Level, Trial Division, Post-Standing Order, Excludes Bond Hearings",
    x = "Division",
    y = "Median Number of Hearings"
  ) +
  scale_y_continuous(breaks = c(0, 1, 2)) +
  theme_minimal()
```

```{r}
# Remove data to tidy environment
rm(hearings_by_division)
rm(hearings_per_case)
rm(hearings_per_case_binned)
```