---
title: "Question 4 (updated)"
author: "Kari Zhou"
date: "2026-04-28"
output: html_document
---
# data import 
```{r}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(tidyr)
library(dplyr)
library(lubridate)
library(ggplot2)
library(stringr)
library(zoo)
getwd()
offense_mapping = read.csv("/Users/karizhou/Desktop/SPRING 26/Capstone/Data/Offense Mapping Data.csv")
cases_charged <- read.csv("/Users/karizhou/Desktop/SPRING 26/Capstone/Data/Cases Charged Data.csv")
diversion_approval <- read.csv("/Users/karizhou/Desktop/SPRING 26/Capstone/Data/diversion_data_approval.csv")

major_crime_levels = c("Homicide/Gangs", "FVPU")
special_labels = c("DVSA", "CCU", "CAEPAD", "HTIC")

duration_colors <- c(
  "less than 45 days" = "wheat",   # darkblue
  "45-90 days"        = "#daa520",
  "3-6 months"        = "orange",
  "6-9 months"        = "tan4",
  "9-12 months"       = "slategrey",
  "1-2 years"         =  "cornflowerblue" ,
  "2-3 years"         = "#4169B0", 
  "3-5 years"         = "darkblue"    # dark golden
)
 

duration_function <- function(dataset, date_received_column, date_final_column)
{
  dataset %>% 
    group_by(case_nbr, party_id) %>% 
    summarise(
      start_time  = min({{ date_received_column }}, na.rm = TRUE),
      end_time    = max({{ date_final_column }},    na.rm = TRUE),
      duration    = as.numeric(difftime(max({{ date_final_column }}), 
                                        min({{ date_received_column }}))),
      .groups = "drop"
    ) 
}
```
# Data cleaning
# might use for later
```{r}
CR_case_data <- read.csv("data/case_data_cr.csv")
CR_case_data <- CR_case_data %>% 
  mutate(
         case_nbr = Case.Number,
         party_id = Party.ID,
         DA_id = DA.ID,
         defendant = Defendant,
         offense_date = as.Date(ifelse(Offense.Date == "" | is.na(Offense.Date),
                                       NA, Offense.Date), 
                                format = "%m/%d/%Y"),
         arrest_date = as.Date(ifelse(Arrest.Date == "" | is.na(Arrest.Date),
                                       NA, Arrest.Date), 
                                format = "%m/%d/%Y"),
         received_date = as.Date(ifelse(Received.Date == "" | is.na(Received.Date),
                                       NA, Received.Date), 
                                format = "%m/%d/%Y"),
         case_status_date = as.Date(ifelse(Case.Status.Date == "" | is.na(Case.Status.Date),
                                       NA, Case.Status.Date), 
                                format = "%m/%d/%Y"),
         offense_code = Offense.Code
         ) %>% 
  select(case_nbr, party_id, DA_id, defendant, offense_date, arrest_date, received_date
         , case_status_date, Offense, offense_code, Unit)
colnames(CR_case_data)
```

```{r}
d_case_data = read.csv("/Users/karizhou/Desktop/SPRING 26/Capstone/Data/case_data_d.csv")
d_case_data <- d_case_data %>% 
  mutate(
         case_nbr = Case.Number,
         party_id = Praty.ID,
         defendant = Defendant,
         arrest_date = as.Date(ifelse(Arrest.Date == "" | is.na(Arrest.Date),
                                       NA, Arrest.Date), 
                                format = "%m/%d/%Y"),
         received_date = as.Date(ifelse(Received.Date == "" | is.na(Received.Date),
                                       NA, Received.Date), 
                                format = "%m/%d/%Y"),
         date_filing = as.Date(ifelse(Case.Status.Date == "" | is.na(Case.Status.Date),
                                       NA, Case.Status.Date), 
                                format = "%m/%d/%Y"),
         offense_code = Offense.Code
         ) %>% 
  select(case_nbr, party_id, defendant, arrest_date, received_date
         , date_filing, Offense, offense_code, Unit)

d_case_data <- d_case_data %>% 
  mutate(date_proxy = case_when(
    is.na(arrest_date) ~ received_date,
    TRUE ~ arrest_date
  )) %>% 
  group_by(case_nbr, party_id) %>% 
  slice(which.min(date_proxy)) %>% 
  select(case_nbr, party_id,defendant, received_date, arrest_date,date_proxy, date_filing, Offense, offense_code, Unit ) %>% 
  filter(received_date >= "2021-06-01",
         date_filing >= "2021-06-01") %>% 
  mutate(case_duration =date_filing - date_proxy) # restrict to only post-covid

d_case_data
```

# cases_charged: used for date information
```{r}
# clean and wrangle cases data, to retrieve filing date
cases_charged <- cases_charged %>% 
   mutate(case_nbr = Case.Number,
         clerk_id = Clerk.ID,
         DA_id = DA.ID,
         party_id = Party.ID,
         offense_code = Offense.Code,
         date_offense = as.Date(ifelse(Offense.Date == "" | is.na(Offense.Date),
                                       NA, Offense.Date), 
                                format = "%m/%d/%y"),
         date_arrest = as.Date(ifelse(Arrest.Date == "" | is.na(Arrest.Date), 
                                      NA, Arrest.Date), 
                                format = "%m/%d/%y"),
         date_received = as.Date(ifelse(Received.Date == ""|is.na(Received.Date),
                                        NA, Received.Date), 
                                format = "%m/%d/%y"),
         date_filing = as.Date(ifelse(Event.Date == ""|is.na(Event.Date),
                                        NA, Event.Date), 
                                format = "%m/%d/%y")
         ) %>% 
  select(case_nbr, clerk_id, DA_id, Defendant, party_id, offense_code,
         Offense, Unit, date_offense, 
         date_arrest, date_received, date_filing) #119723

colnames(cases_charged)
```

# cases_closed: might use for cases without filing date 
```{r}
cases_closed <- read.csv("/Users/karizhou/Desktop/SPRING 26/Capstone/Data/Cases Closed Data.csv")

cases_closed <- cases_closed %>% 
   mutate(case_nbr = Case.Number,
         clerk_id = Clerk.ID,
         DA_id = DA.ID,
         party_id = Party.ID,
         offense_code = Offense.Code,
         date_arrest = as.Date(ifelse(Arrest.Date == "" | is.na(Arrest.Date), 
                                      NA, Arrest.Date), 
                                format = "%m/%d/%y"),
         date_received = as.Date(ifelse(Received.Date == ""|is.na(Received.Date),
                                        NA, Received.Date), 
                                format = "%m/%d/%y"),
         date_filing = as.Date(ifelse(Event.Date == ""|is.na(Event.Date),
                                        NA, Event.Date), 
                                format = "%m/%d/%y")
         ) %>% 
  select(case_nbr, clerk_id, DA_id, Defendant, party_id, offense_code,
         Offense, Unit, 
         date_arrest, date_received, date_filing) #119723


cases_closed <- cases_closed %>% 
  mutate(date_proxy = case_when(
    is.na(date_arrest) ~ date_received,
    TRUE ~ date_arrest
  )) %>% 
  group_by(case_nbr, party_id) %>% 
  slice(which.min(date_proxy)) %>% 
  select(case_nbr, party_id,Defendant, date_received,date_proxy, date_filing, Offense, offense_code, Unit ) %>% 
  filter(date_received >= "2021-06-01",
         date_filing >= "2021-06-01") %>% 
  mutate(case_duration =date_filing - date_proxy) # restrict to only post-covid

colnames(cases_closed)
```

# Exclude special unit cases that dont qualify for diversion (typically)
```{r}
cases_charged <- cases_charged %>% 
  filter(!Unit %in% special_labels,
         !Unit %in% major_crime_levels) %>%  
  # exclude cases belong to special units, because they won't get diversion. 
  group_by(case_nbr, party_id) %>% 
  mutate(date_proxy = case_when(
    date_arrest < date_received ~ date_arrest,
    TRUE ~ date_received
  )) %>% 
  group_by(case_nbr,party_id) %>% 
  slice(which.min(date_proxy),
        which.max(date_filing)) %>% 
  mutate(case_duration = date_filing - date_proxy) #43502

cases_charged = cases_charged[!duplicated(cases_charged),] # clear duplicates 

cases_charged <- cases_charged %>% filter(case_duration >=0) #21880
colnames(cases_charged)
# Cases_charged will be used for date calculation. 
```


```{r}
cases_charged %>% filter(is.na(date_filing)) # no NA
cases_charged %>% filter(is.na(date_arrest)) 
```

```{r}
diversion_approval <- diversion_approval %>% 
   mutate(case_nbr = Case.Number,
         DA_id = DA.ID,
         party_id = Party.ID,
         date_approval = as.Date(ifelse(Event.Date == "" | is.na(Event.Date),
                                       NA, Event.Date), 
                                format = "%m/%d/%Y")) %>% 
  select(,-c(Case.Number, DA.ID, Party.ID, Event.Date)) #3168

# exclude tester rows
diversion_approval <- diversion_approval %>% 
 filter(case_nbr != "datest123") #3166
```


# retrieve dates from cases charged
```{r}
diversion_approval <- left_join(diversion_approval, cases_charged[,c("case_nbr","party_id", "offense_code","Offense",
                                                                  "Unit", "date_proxy", "date_filing","case_duration")],
                                by = c("case_nbr", "party_id")) 

diversion_approval = diversion_approval[!duplicated(diversion_approval),] # 3167, with multiple charges with each defendant

```

```{r}
diversion_nofiling <- diversion_approval %>% 
  filter(is.na(date_filing),
         date_approval >= "2021-06-01")
  # 1228
diversion_nofiling
```

```{r}
diversion_approval %>% group_by(party_id, case_nbr) %>% 
  summarise(count = n()) %>% 
  filter(count >1)

# on case defendant level, keep the earlier date_approval
diversion_approval <- diversion_approval %>% 
  group_by(party_id, case_nbr) %>% 
  slice(which.min(date_approval)) #2816

# check again
diversion_approval %>% group_by(party_id, case_nbr) %>% 
  summarise(count = n()) %>% 
  filter( count >1) # 0

diversion_approval <- diversion_approval[!duplicated(diversion_approval),] #2816 
diversion_approval <- diversion_approval %>% 
  filter(!is.na(case_duration),
         date_proxy >= "2021-06-01") #504
```

```{r}
diversion_nofiling %>% 
  mutate(category = case_when(
    startsWith(case_nbr, "D") ~ "Pre-indictment",
    TRUE ~ "Post-Indictment"
  )) %>% 
  group_by(category) %>% 
  summarise(count = n()) %>% 
  ggplot(aes(x = category, y = count))+
  geom_col(fill = "navy")+
  labs(x = "Status in Judicial Process",
       y = "Number of cases",
       title = "Diversion cases without filing date and their status")+
  theme_bw()
```
```{r}
head(d_case_data)
```

```{r}
# left join with d case data, to find a proxy filing date for all cases that weren't indicted. 
diversion_preindict <- left_join(diversion_nofiling[, c(1:6)], d_case_data[,c("case_nbr", "party_id", "offense_code", "Offense", "Unit", 
                                                                   "date_proxy", "date_filing", 
                                                                   "case_duration")], 
                                by = c("case_nbr", "party_id"))

diversion_preindict %>% filter(is.na(date_filing)) #296/1228 is NA 
```

```{r}
# left join with closed cases data, to find a proxy filing date for all cases that were closed. 
diversion_closed <- left_join(diversion_nofiling[, c(1:6)], cases_closed[,c("case_nbr", "party_id", "offense_code", "Offense", "Unit", 
                                                                   "date_proxy", "date_filing", 
                                                                   "case_duration")], 
                                by = c("case_nbr", "party_id"))
diversion_closed %>% 
  filter(is.na(date_filing)) #868/1228 is NA
```


```{r}
#there could be overlaps, so we combine the two datasets first
diversion_combined = rbind(diversion_closed, diversion_preindict)

# then, for the same case-defendant approval, we only keep the earliest approval and the longer duration
# (might be approved by diff units twice)
 diversion_combined = diversion_combined %>% 
  filter(!is.na(date_filing)) %>% 
  group_by(case_nbr, party_id) %>% 
  slice(which.min(date_approval),
        which.max(case_duration))
# check for duplicates
diversion_combined <- diversion_combined[!duplicated(diversion_combined),] 
diversion_combined #925
```

```{r}
# preindictment-cases
preindictment_diversion <- diversion_combined %>% 
  filter(startsWith(case_nbr, "D")) #909

# post-indictment closed cases
postindictment_diversion <- diversion_combined %>% 
  filter(!startsWith(case_nbr, "D")) #16
```

```{r}
approval_analysis <- diversion_approval%>% 
  mutate(year = year(date_proxy),
         month= month(date_proxy),
         quarter = as.yearqtr(date_proxy, format = "%Y-%m-%d")) %>% 
  filter(quarter >="2021Q2") %>% 
  group_by(year, month) %>% 
  summarise(average = mean(case_duration,2)) %>% 
   mutate(duration = as.numeric(average),
          category = case_when(
            duration <= 45 ~ "less than 45 days" ,
            duration > 45 & duration <=90 ~ "45-90 days",
            duration > 90 & duration <= 180 ~ "3-6 months",
            duration > 180 & duration < 270 ~ "6-9 months",
            duration >= 270 & duration <= 365 ~ "9-12 months",
            duration > 365 & duration < 730 ~ "1-2 years",
            duration >= 730 &duration < 1095 ~ "2-3 years",
            duration >= 1095 & duration < 1825 ~ "3-5 years"
          )) %>% 
  mutate(
  duration_category = factor(category, levels = c(
    "less than 45 days",
    "45-90 days",
    "3-6 months",
    "6-9 months",
    "9-12 months",
    "1-2 years",
    "2-3 years",
    "3-5 years"
  )))

diversion_graph = ggplot(approval_analysis, 
       aes(x = factor(month), 
               y = duration, 
               fill = duration_category)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = duration_colors) +
 facet_wrap(~ year, strip.position = "bottom") +        # one panel per year
  scale_x_discrete(
  breaks = c(1, 7, 12),
  labels = c("Jan", "July", "Dec")
)+
  ylim(0, 1500)+
  labs(x = "Time started", y = "Average Duration (days)", legend= "Duration",
       title = "Average Duration for Cases Approved for Diversion (Monthly)",
       )+
  theme_bw()
diversion_graph
```



```{r}
# SVC:
svc <- c("ARMED ROBBERY", "KIDNAPPING", "AGGRAVATED ASSAULT", "AGGRAVATED BATTERY", "TRAFFICKING","	
HIJACKING A MOTOR VEHICLE")

svf <- c("MURDER - FELONY", "RAPE", "AGGRAVATED CHILD MOLESTATION", "AGGRAVATED SODOMY",
         "AGGRAVATED SEXUAL BATTERY")
 
# set up a parallel comparsion for cases could qualify diversion 
cases_non_diversion <- cases_charged %>% filter(!Offense %in% svf,
                                                !Offense %in% svc,
                                                !startsWith(Offense, "HIJACKING A MOTOR VEHICLE"),
                                                !startsWith(Offense, "TRAFFICKING"),
                                                !startsWith(Offense, c("AGGRAVATED ASSAULT")),
                                                !startsWith(Offense, "AGGRAVATED BATTERY"),
                                                !Unit %in% special_labels,
                                                !Unit %in% major_crime_levels) #21466

cases_non_diversion$party_id <- as.character(cases_non_diversion$party_id)
cases_non_diversion <- cases_non_diversion %>% 
  group_by(party_id) %>% 
  slice(which.min(date_offense)) %>% 
  filter(date_proxy >= "2021-06-01")#4962
```

```{r}
cases_non_diversion_analysis = cases_non_diversion %>% 
  mutate(year = year(date_proxy),
         month= month(date_proxy),
         quarter = as.yearqtr(date_proxy, format = "%Y-%m-%d")) %>% 
 filter(quarter >="2021Q2") %>% 
  group_by(year, month) %>% 
  summarise(average = mean(case_duration,2)) %>% 
   mutate(duration = as.numeric(average),
          category = case_when(
            duration <= 45 ~ "less than 45 days" ,
            duration > 45 & duration <=90 ~ "45-90 days",
            duration > 90 & duration <= 180 ~ "3-6 months",
            duration > 180 & duration < 270 ~ "6-9 months",
            duration >= 270 & duration <= 365 ~ "9-12 months",
            duration > 365 & duration < 730 ~ "1-2 years",
            duration >= 730 &duration < 1095 ~ "2-3 years",
            duration >= 1095 & duration < 1825 ~ "3-5 years"
          )) %>% 
  mutate(
  duration_category = factor(category, levels = c(
    "less than 45 days",
    "45-90 days",
    "3-6 months",
    "6-9 months",
    "9-12 months",
    "1-2 years",
    "2-3 years",
    "3-5 years"
  )))

non_diversion_graph = ggplot(cases_non_diversion_analysis, 
       aes(x = factor(month), 
               y = duration, 
               fill = duration_category)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = duration_colors) +
 facet_wrap(~ year, strip.position = "bottom") +        # one panel per year
  scale_x_discrete(
  breaks = c(1, 7, 12),
  labels = c("Jan", "July", "Dec")
)+
  ylim(0, 1500)+
  labs(x = "Time", y = "Average Duration (days)", legend= "Duration",
       title = "Average Duration for Cases without Diversion Enrollment (Monthly)",
       )+
  theme_bw()

non_diversion_graph
```

```{r}
cases_non_diversion_anaysis <- cases_non_diversion %>% 
  select(case_duration)

non_diversion_mean = mean(cases_non_diversion_anaysis$case_duration)
non_diversion_mean
```

```{r}
diversion_analysis <- diversion_approval %>% 
  select(case_duration )

diversion_analysis_mean = mean(diversion_analysis$case_duration)
diversion_analysis_mean
```


```{r}
preindicted_analysis <- preindictment_diversion %>% 
  select(case_duration)

preindicted_analysis_mean = mean(preindicted_analysis$case_duration)
preindicted_analysis_mean
```
```{r}
post_analysis <- postindictment_diversion %>% 
  select(case_duration)

post_analysis_mean = mean(post_analysis$case_duration)
post_analysis_mean

```

```{r}
analysis = data.frame(group = c("Diversion, Charged", "Non-Diversion, Charged", 
                                "Diversion, Not Indicted", "Diversion, Indicted but Closed"),
                         values = c(diversion_analysis_mean, non_diversion_mean,
                                    preindicted_analysis_mean, post_analysis_mean))

class(analysis$values)
```
# ai use:
*prompt" *
for  "difftime" variable,i wnat to transform the unit from days to year: 
*answer:*
You can convert a difftime variable from days to years like this:
df$difftime_years <- as.numeric(df$difftime, units = "days") / 365.25

```{r}

analysis$values_years <- as.numeric(analysis$values, units = "days") / 365.25

```

```{r}
ggplot(analysis, aes(group, values_years, fill = group))+
  geom_col()+
  scale_fill_manual(values = c("Diversion, Charged" = "tan4",
                      "Non-Diversion, Charged"="darkblue",
                      "Diversion, Not Indicted" = "red4",
                      "Diversion, Indicted but Closed" = "lightblue" ))+
  theme_bw()+
  labs(x= "Diversion Enrollment status",
       y = "Time taken (Year)",
       title = "Average Duration for Diversion vs. Non-Diversion Cases (Year), June 2021 - Oct. 2025")

```

```{r}
analysis_simple = analysis %>% filter(group == c("Diversion, Charged", "Non-Diversion, Charged"))
ggplot(analysis_simple, 
       aes(group, values_years, fill = group))+
  geom_col()+
  scale_fill_manual(values = c("Diversion, Charged" = "tan4",
                      "Non-Diversion, Charged"="darkblue"))+
                     # "Diversion, Not Indicted" = "red4",s
                      #"Diversion, Indicted but Closed" = "lightblue" ))+
  theme_bw()+
  labs(x= "Diversion Enrollment status",
       y = "Time taken (Year)",
       title = "Average Duration for Diversion vs. Non-Diversion Cases (Year), June 2021-Oct. 2025")
```

```{r}
diversion_analysis$party_id <- as.character(diversion_analysis$party_id
                                            )
stats_comp =  rbind(diversion_analysis %>% mutate(group = "diversion"),
      cases_non_diversion_anaysis %>% mutate(group = "non-Diversion")) %>% ungroup()%>% select(case_duration, group)

stats_comp %>%
  summarise(p_value = round(t.test(case_duration ~ group)$p.value))

t.test(case_duration ~ group, data = stats_comp)
```


